---
title: "Community_trajectory"
author: "Joshua G. Smith"
date: '2022-06-16'
output: html_document
---



#load packages
```{r}
require(tidyverse)
require(stringr)
require(janitor)
require(vegan)
require(ggplot2)
library(parallel)
library(scales)
library(here)
require(ggpubr)
```


#load data

```{r}

data_path <- "/home/shares/ca-mpa/data/sync-data/processed_data/ecological_community_data/year_level"


#load CCFRP
input_file <- "CCFRP_mpa_year.csv" 
CCFRP_counts <- read.csv(file.path(data_path, input_file))%>%
                filter(region4=='central')
                     
              

#load kelp upc
input_file <- "kelp_upc_mpa_year.csv" 
kelp_upc_counts <- read.csv(file.path(data_path, input_file))%>%
                    filter(region4=='central',
                           mpa_defacto_designation=='smr'|mpa_defacto_designation=='ref')

#load kelp swath
input_file <- "kelp_swath_mpa_year.csv" 
kelp_swath_counts <- read.csv(file.path(data_path, input_file))%>%
                    filter(region4=='central')
#load kelp_fish
input_file <- "kelp_fish_mpa_year.csv" 
kelp_fish_counts <- read.csv(file.path(data_path, input_file))%>%
                    filter(region4=='central')

#load deep reef
input_file <- "deep_reef_mpa_year.csv" 
deep_reef_counts <- read.csv(file.path(data_path, input_file))%>%
                    filter(region4=='central')

#load rocky intertidal
input_file <- "rocky_mpa_year.csv" 
rocky_counts <- read.csv(file.path(data_path, input_file))%>%
                    filter(region4=='central')

```


#OLD -- test ordinate across YEARS and regions
```{r, eval=FALSE}

# CCFRP_test <- CCFRP 
# 
# year <- as.numeric(CCFRP_test$year)
# region <- CCFRP_test$region4
# mpa_designation <- CCFRP_test$mpa_designation
# 
# CCFRP_test <- CCFRP_test%>%
#               dplyr::select(10:ncol(.))
# 
# CCFRP_relative <- decostand(CCFRP_test, method = "total") #convert absolute abundance to relative abundance
#               
# distmat <- 
#     vegdist(CCFRP_relative, method = "eucl") #generates a euclidean based distance matrix
# 
# 
# ord <- metaMDS(distmat, distance="eucl") #ordinates the dist matrix
# 
# 
# 
# plot(ord)
# ordiellipse(ord, year)
# ordispider(ord, year, label=F)
# 
# 
# 
# scrs <- scores(ord)
# 
# scrs <- cbind(as.data.frame(scrs), year, region, mpa_designation) #to facilitate computing centroids, add group var
# scrs$region <- factor(scrs$region, level = c("north","central","north islands","south")) #set order
# scrs$mpa_designation <- factor(scrs$mpa_designation, level = c("smr","ref")) #set order
# 
# 
# cent <- aggregate(cbind(NMDS1, NMDS2) ~ year+region+mpa_designation, data = scrs, FUN = mean) #computes centroids by MHW and region
# 
# segs <- merge(scrs, setNames(cent, c('year','oNMDS1','oNMDS2')),
#               by = 'year', sort=FALSE) #to draw the spider, we need to use geom_segment() which required coordinates to draw the segment from and to. Our 'to'
#                                        # coordinates, the xend and yend aesthetics will be the centroids. So we need to replicate the group centroid for each
#                                        # observation in the group. T
# 
# 
# 
# 
# ggplot(scrs, aes(x = NMDS1, y=NMDS2, shape = region, color = mpa_designation, group=region)) +
#   #geom_segment(data = segs,
#                #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
#   geom_point(data = cent, size=5) +                           #centroids
#   geom_path(data = cent, size = 0.5, #aes(color=year)
#             )+           #line connecting groups (years)
#   #geom_point() +                                              #sample scores
#   #coord_fixed() +
#   geom_text(data=cent, aes(label=year), position=position_jitter(width=0.02, height=0.01))+
#   #xlim(-4.5,4.5)+
#   #ylim(-0.45,0.65)+
#   facet_wrap(~mpa_designation,
#              labeller = labeller(mpa_designation =
#                                    c("ref"="REF",
#                                      "smr"="SMR")))+
#   ggtitle("CCFRP assemblage structure over time")+
#   theme(strip.text = element_text(size=18, face="bold"))+
#   theme(plot.title = element_text(size=20, face="bold"))+
#   theme(text = element_text(20))
#   
#   #theme(legend.position='none')
# 

```





#clean up data for ordination
```{r}
#CCFRP processing--------------------------------------------------------------
CCFRP_process <- CCFRP_counts %>%
                  rowwise() %>%
                  dplyr::mutate(sum = sum(across(10:ncol(.)), na.rm = T)) %>%
                  filter(!(sum==0))%>%
                  dplyr::select(-sum)%>%
                  mutate(desig_state = paste(mpa_designation,MHW))%>%
                  dplyr::select(desig_state, everything())#remove rows containing only zeros

#define grouping vars
CCFRP_group_vars <- CCFRP_process %>%
                    dplyr::select(1:9)

#define data for ordination
CCFRP_ord_data <- CCFRP_process %>%
              dplyr::select(10:ncol(.))


#calculate relative abundance
CCFRP_rel <- decostand(CCFRP_ord_data, method = "hellinger") %>%
            dplyr::select(where(~any(. !=0)))

#generate a BC dissim matrix
CCFRP_distmat <- vegdist(CCFRP_rel, method = "bray", na.rm=T) 



#kelp swath processing---------------------------------------------------------
kelp_swath <- kelp_swath_counts %>%
                  rowwise() %>%
                  dplyr::mutate(sum = sum(across(8:ncol(.), na.rm = T))) %>%
                  filter(!(sum==0))%>% #remove rows containing only zeros
                  dplyr::select(-sum, unidentified_mobile_invert_species, 
                                no_organisms_present_in_this_sample)%>%
                  mutate(desig_state = paste(mpa_defacto_designation,MHW))%>%
                  dplyr::select(desig_state, everything())%>%
                  filter(mpa_defacto_designation=="smr" | mpa_defacto_designation=="ref")#remove rows containing only zeros #drop sum columns and non-species categories

#define grouping vars
kelp_swath_group_vars <- kelp_swath%>%
                    dplyr::select(1:8)

#define data for ordination
kelp_swath_ord_data <- kelp_swath%>%
              ungroup()%>%
              dplyr::select(9:ncol(.))
                         #%>%    #remove all-zero columns
              #mutate_if(is.character, as.numeric)
           

          
#calculate relative abundance
kelp_swath_rel <- decostand(kelp_swath_ord_data, method = "hellinger")
  
#generate a BC dissim matrix
kelp_swath_distmat <- 
    vegdist(kelp_swath_rel, method = "bray", na.rm=T) #generates a BC dissim matrix




#kelp upc processing---------------------------------------------------------

kelp_upc <- kelp_upc_counts %>%
                  rowwise() %>%
                  dplyr::mutate(sum = sum(across(8:ncol(.), na.rm = T))) %>%
                  #filter(!(sum==0))%>% #remove rows containing only zeros
                  dplyr::select(-sum, bare_rock,unidentified_fish,
                                bare_sand, shell_debris)%>%
                  mutate(desig_state = paste(mpa_defacto_designation,MHW))%>%
                  dplyr::select(desig_state, everything())%>%
                  filter(mpa_defacto_designation=="smr" | mpa_defacto_designation=="ref")#drop sum columns and non-species categories
              

kelp_upc[is.na(kelp_upc)] = 0                  

#define grouping vars
kelp_upc_group_vars <- kelp_upc%>%
                    dplyr::select(1:8)

#define data for ordination
kelp_upc_ord_data <- kelp_upc%>%
              ungroup()%>%
              dplyr::select(9:ncol(.))


#calculate relative abundance
kelp_upc_rel <- decostand(kelp_upc_ord_data, method = "hellinger")

#generate a BC dissim matrix
kelp_upc_distmat <- 
    vegdist(kelp_upc_rel, method = "bray", na.rm=T) #generates a BC dissim matrix




#kelp fish processing---------------------------------------------------------

kelp_fish <- kelp_fish_counts %>%
                  rowwise() %>%
                  dplyr::mutate(sum = sum(across(8:ncol(.), na.rm = T))) %>%
                  filter(!(sum==0))%>% #remove rows containing only zeros
                  dplyr::select(-sum) %>%
                               dplyr:: select(where(~ any(. != 0)))%>%
                  mutate(desig_state = paste(mpa_defacto_designation,MHW))%>%
                  dplyr::select(desig_state, everything())%>%
                  filter(mpa_defacto_designation=="smr" | mpa_defacto_designation=="ref")#drop sum columns and non-species categories  #drop sum columns and non-species categories
             

#define grouping vars
kelp_fish_group_vars <- kelp_fish %>%
                    dplyr::select(1:8)

#define data for ordination
kelp_fish_ord_data <- kelp_fish %>%
              ungroup() %>%
              dplyr::select(9:ncol(.))

#calculate relative abundance
kelp_fish_rel <- decostand(kelp_fish_ord_data, method = "hellinger")

#kelp_fish_rel <- kelp_fish_rel %>% slice_sample(n=2000) # testing if work on smaller

#generate a BC dissim matrix
kelp_fish_distmat <- 
    vegdist(kelp_fish_rel, method = "bray", na.rm=T) #generates a BC dissim matrix



#deep reef processing---------------------------------------------------------

deep_reef <- deep_reef_counts %>%
                  rowwise() %>%
                  dplyr::mutate(sum = sum(across(8:ncol(.), na.rm = T))) %>%
                  filter(!(sum==0))%>% #remove rows containing only zeros
                  dplyr::select(-c(sum, schooling_10_15_cm_sebastes_sp,
                                schooling_10_15_cm_sebastes_sp,
                                young_of_year_10_cm_sebastes_sp,
                                synodus_lucioceps_or_ophiodon_elongatus,
                                sebastes_melanops_or_mystinus_or_diaconus))%>%
                  mutate(desig_state = paste(mpa_defacto_designation,MHW))%>%
                  dplyr::select(desig_state, everything())%>%
                  filter(mpa_defacto_designation=="smr" | mpa_defacto_designation=="ref")#drop sum columns and non-species categories

#define grouping vars
deep_reef_group_vars <- deep_reef%>%
                    dplyr::select(1:8)

#define data for ordination
deep_reef_ord_data <- deep_reef%>%
              ungroup()%>%
              dplyr::select(9:ncol(.))

#calculate relative abundance
deep_reef_rel <- decostand(deep_reef_ord_data, method="hellinger")

#generate a BC dissim mat
deep_reef_distmat <- 
    vegdist(deep_reef_rel, method = "bray", na.rm=T) #generates a BC dissim matrix



#Intertidal processing---------------------------------------------------------

rocky_counts <- rocky_counts %>%
                  mutate(MHW = ifelse(year>=2014 & year<=2016, "during",ifelse(year<2014, "before","after")))%>%
                  mutate(desig_state = paste(mpa_designation,MHW))%>%
                  dplyr::select(desig_state, MHW, everything())%>%
                  filter(mpa_designation=="smr" | mpa_designation=="ref")

#define grouping vars
rocky_group_vars <- rocky_counts%>%
                    dplyr::select(1:9)

#define data for ordination
rocky_ord_data <- rocky_counts %>%
              ungroup() %>%
              dplyr::select(10:ncol(.))

#calculate relative abundance
rocky_rel <- decostand(rocky_ord_data, method = "hellinger")

#generate a BC dissim matrix
rocky_distmat <- 
    vegdist(rocky_rel, method = "bray", na.rm=T) #generates a BC dissim matrix





#save group vars
#data_path <- "/home/shares/ca-mpa/data/sync-data/processed_data/ecological_community_data/year_level"
#save(CCFRP_group_vars,kelp_swath_group_vars,kelp_upc_group_vars,kelp_fish_group_vars,
#     deep_reef_group_vars,rocky_group_vars,
#     file = file.path(data_path, "group_vars.rda"))


#save distance matrices
#data_path <- "/home/shares/ca-mpa/data/sync-data/processed_data/ecological_community_data/year_level"
#save(CCFRP_distmat,kelp_swath_distmat,kelp_upc_distmat,kelp_fish_distmat,
#     deep_reef_distmat,rocky_distmat,
#     file = file.path(data_path, "distance_matrices_BC.rda"))



```


## ordinate all datasets. NOTE: THIS CHUNK TAKES A LOT OF COMPUTATION TIME. Resulting objects are exported as .Rdata
#and can be loaded directly in next chunk.

```{R}
# setting the seed 
set.seed(123)
num_cores=20

message("Processing CCFRP_ord")
CCFRP_ord <- metaMDS(CCFRP_distmat, distance="bray", trymax = 100, parallel = num_cores) #ordinates the dist matrix
#CCFRP_ord <- metaMDS(CCFRP_distmat, previous.best = CCFRP_ord, noscore=0.2, trymax=100)
message("Done processing CCFRP_ord")

message("Processing kelp_swath_ord")
kelp_swath_ord <- metaMDS(kelp_swath_distmat, distance="bray", trymax = 100, parallel = num_cores) #ordinates the dist matrix
message("Done processing kelp_swath_ord")

message("Processing kelp_fish_ord")
kelp_fish_ord <- metaMDS(kelp_fish_distmat, distance="bray", noshare=0.2, parallel = num_cores)
kelp_fish_ord <- metaMDS(kelp_fish_distmat, previous.best = kelp_fish_ord, noshare=0.1, trymax=300, parallel = num_cores)
message("Done processing kelp_fish_ord")

message("Processing kelp_upc_ord")
kelp_upc_ord <- metaMDS(kelp_upc_distmat, distance="bray", noshare=0.2, parallel = num_cores)
kelp_upc_ord <- metaMDS(kelp_upc_distmat,previous.best=kelp_upc_ord, noshare=0.2, trymax = 300, parallel = num_cores)
message("Done processing kelp_upc_ord")

message("Processing deep_reef_ord")
deep_reef_ord <- metaMDS(deep_reef_distmat, distance="bray", trymax = 100, parallel = num_cores)
message("Done processing deep_reef_ord")

message("Processing rocky_ord")
rocky_ord <- metaMDS(rocky_distmat, distance="bray", trymax = 100, parallel = num_cores)
message("Done processing rocky_ord")


#export to .RData
data_path <- "/home/shares/ca-mpa/data/sync-data/processed_data/ecological_community_data/year_level"
save(CCFRP_ord, kelp_swath_ord, kelp_fish_ord, kelp_upc_ord, deep_reef_ord, rocky_ord, file = file.path(data_path, "bray_nmds_scores.rda"))

```



#CCFRP scores and plot
```{r}

data_path <- "/home/shares/ca-mpa/data/sync-data/processed_data/ecological_community_data/"
input_file <- "bray_nmds_scores.rda"

data <- load(file.path(data_path, input_file))



#plot(ord)
#ordiellipse(ord, year)
#ordispider(ord, year, label=F)

#CCFRP_groups <- CCFRP_groups %>%
                # mutate(MHW = ifelse(year>=2014 & year<=2016, "during",ifelse(year<2014, "before","after")))

#en = envfit(CCFRP_ord, CCFRP_groups, permutations=999, na.rm=TRUE)


scrs <- scores(CCFRP_ord)

scrs <- cbind(as.data.frame(scrs), desig_state=CCFRP_group_vars$desig_state, year=CCFRP_group_vars$year, MHW=CCFRP_group_vars$MHW, region4=CCFRP_group_vars$region4, mpa_designation=CCFRP_group_vars$mpa_designation) #to facilitate computing centroids, add group var

scrs$region4 <- factor(scrs$region4, level = c("north","central","north islands","south")) #set order
scrs$MHW <- factor(scrs$MHW, level = c("before","during","after")) #set order
scrs$mpa_designation <- factor(scrs$mpa_designation, level = c("smr","ref")) #set order


cent <- aggregate(cbind(NMDS1, NMDS2) ~ MHW+region4+mpa_designation, data = scrs, FUN = mean) #computes centroids by MHW and region

segs <- merge(scrs, setNames(cent, c('year','oNMDS1','oNMDS2')),
              by = 'year', sort=FALSE) #to draw the spider, we need to use geom_segment() which required coordinates to draw the segment from and to. Our 'to'
                                       # coordinates, the xend and yend aesthetics will be the centroids. So we need to replicate the group centroid for each
                                       # observation in the group. T




ggplot(scrs, aes(x = NMDS1, y=NMDS2,  color = mpa_designation)) +
  #geom_segment(data = segs,
               #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
  geom_point(data = cent, size=5, aes(shape=MHW)) +                           #centroids
  geom_path(data = cent, size = 0.5, #aes(color=year)
            )+           #line connecting groups (years)
  #geom_point() +                                              #sample scores
  #coord_fixed() +
  #geom_text(data=cent, aes(label=year), position=position_jitter(width=0.02, height=0.01))+
  #xlim(-4.5,4.5)+
  #ylim(-0.45,0.65)+
  #facet_wrap(~mpa_designation,
  #           labeller = labeller(mpa_designation =
  #                                 c("ref"="REF",
  #                                   "smr"="SMR")))+
  ggtitle("CCFRP assemblage structure over time")+
  theme(strip.text = element_text(size=18, face="bold"))+
  theme(plot.title = element_text(size=20, face="bold"))+
  theme(text = element_text(20)) +
  stat_ellipse(type="t")


ord_scores <- cent %>%
              dplyr:: select(NMDS1, NMDS2, groups=MHW)





NMDS = data.frame(MDS1 = CCFRP_ord$points[,1], MDS2 = CCFRP_ord$points[,2],group=CCFRP_group_vars$desig_state, mpa_desig=CCFRP_group_vars$mpa_designation)

plot(CCFRP_ord)
ord<- ordiellipse(CCFRP_ord, CCFRP_group_vars$desig_state, kind = "se",conf=0.95, label=T)

NMDS$group <- as.factor(NMDS$group)

df_ell <- data.frame()
for(g in levels(NMDS$group)){
  df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$group==g,],
                  vegan:::veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale)))
                                ,group=g))
}

df_ell$mpa_type <- word(df_ell$group, 1)

#ggplot(data = NMDS, aes(MDS1, MDS2)) + geom_point(aes(color = group)) +
#  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=group), size=1, linetype=2)


#reorder for plotting
cent$mpa_designation <- factor(cent$mpa_designation, level=c("smr","ref"))
cent$MHW <- factor(cent$MHW, level=c("before","during","after"))




CCFRP<- ggplot(data=cent, aes(NMDS1, NMDS2)) +
  #geom_segment(data = segs,
               #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
  geom_point(aes(shape=MHW, color=mpa_designation), size=4) +                           #centroids
  geom_path(data = df_ell, aes(NMDS1, NMDS2, colour=mpa_type, group=group), size = 0.5, linetype=2 #aes(color=year)
            )+           #line connecting groups (years)
  #geom_point() +                                              #sample scores
  #coord_fixed() +
  #geom_text(data=cent, aes(label=year), position=position_jitter(width=0.02, height=0.01))+
  #xlim(-4.5,4.5)+
  #ylim(-0.45,0.65)+
  #facet_wrap(~mpa_designation,
  #           labeller = labeller(mpa_designation =
  #                                 c("ref"="REF",
  #                                   "smr"="SMR")))+
  ggtitle("CCFRP")+
  scale_color_manual(values=c('red','blue'))+
  theme(strip.text = element_text(size=18, face="bold"))+
  theme(plot.title = element_text(size=16, face="bold"))+
  theme(text = element_text(20))


CCFRP


```

#Kelp swath scores and plot
```{r}

scrs<- scores(kelp_swath_ord)
scrs <- cbind(as.data.frame(scrs), desig_state=kelp_swath_group_vars$desig_state, year=kelp_swath_group_vars$year, MHW=kelp_swath_group_vars$MHW, region4=kelp_swath_group_vars$region4, mpa_designation=kelp_swath_group_vars$mpa_defacto_designation) #to facilitate computing centroids, add group var

cent <- aggregate(cbind(NMDS1, NMDS2) ~ MHW+region4+mpa_designation, data = scrs, FUN = mean) #computes centroids by MHW and region




NMDS = data.frame(MDS1 = kelp_swath_ord$points[,1], MDS2 = kelp_swath_ord$points[,2],group=kelp_swath_group_vars$desig_state, mpa_desig=kelp_swath_group_vars$mpa_defacto_designation)


plot(kelp_swath_ord)
ord<- ordiellipse(kelp_swath_ord, kelp_swath_group_vars$desig_state, kind = "se",conf=0.95, label=T)

NMDS$group <- as.factor(NMDS$group)

df_ell <- data.frame()
for(g in levels(NMDS$group)){
  df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$group==g,],
                  vegan:::veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale)))
                                ,group=g))
}

df_ell$mpa_type <- word(df_ell$group, 1)

#ggplot(data = NMDS, aes(MDS1, MDS2)) + geom_point(aes(color = group)) +
#  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=group), size=1, linetype=2)


#reorder for plotting
cent$mpa_designation <- factor(cent$mpa_designation, level=c("smr","ref"))
cent$MHW <- factor(cent$MHW, level=c("before","during","after"))

KF_swath <- ggplot(data=cent, aes(NMDS1, NMDS2)) +
  #geom_segment(data = segs,
               #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
  geom_point(aes(shape=MHW, color=mpa_designation), size=4) +                           #centroids
  geom_path(data = df_ell, aes(NMDS1, NMDS2, colour=mpa_type, group=group), size = 0.5, linetype=2 #aes(color=year)
            )+           #line connecting groups (years)
  #geom_point() +                                              #sample scores
  #coord_fixed() +
  #geom_text(data=cent, aes(label=year), position=position_jitter(width=0.02, height=0.01))+
  #xlim(-4.5,4.5)+
  #ylim(-0.45,0.65)+
  #facet_wrap(~mpa_designation,
  #           labeller = labeller(mpa_designation =
  #                                 c("ref"="REF",
  #                                   "smr"="SMR")))+
  ggtitle("kelp forest swath ")+
  scale_color_manual(values=c('red','blue'))+
  theme(strip.text = element_text(size=18, face="bold"))+
  theme(plot.title = element_text(size=16, face="bold"))+
  theme(text = element_text(20))


```


#Kelp upc scores and plot
```{r}

scrs<- scores(kelp_upc_ord)
scrs <- cbind(as.data.frame(scrs), desig_state=kelp_upc_group_vars$desig_state, year=kelp_upc_group_vars$year, MHW=kelp_upc_group_vars$MHW, region4=kelp_upc_group_vars$region4, mpa_designation=kelp_upc_group_vars$mpa_defacto_designation) #to facilitate computing centroids, add group var

cent <- aggregate(cbind(NMDS1, NMDS2) ~ MHW+region4+mpa_designation, data = scrs, FUN = mean) #computes centroids by MHW and region




NMDS = data.frame(MDS1 = kelp_upc_ord$points[,1], MDS2 = kelp_upc_ord$points[,2],group=kelp_upc_group_vars$desig_state, mpa_desig=kelp_upc_group_vars$mpa_defacto_designation)


plot(kelp_upc_ord)
ord<- ordiellipse(kelp_upc_ord, kelp_upc_group_vars$desig_state, kind = "se",conf=0.95, label=T)

NMDS$group <- as.factor(NMDS$group)

df_ell <- data.frame()
for(g in levels(NMDS$group)){
  df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$group==g,],
                  vegan:::veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale)))
                                ,group=g))
}

df_ell$mpa_type <- word(df_ell$group, 1)

#ggplot(data = NMDS, aes(MDS1, MDS2)) + geom_point(aes(color = group)) +
#  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=group), size=1, linetype=2)


#reorder for plotting
cent$mpa_designation <- factor(cent$mpa_designation, level=c("smr","ref"))
cent$MHW <- factor(cent$MHW, level=c("before","during","after"))

(KF_upc <- ggplot(data=cent, aes(NMDS1, NMDS2)) +
  #geom_segment(data = segs,
               #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
  geom_point(aes(shape=MHW, color=mpa_designation), size=4) +                           #centroids
  geom_path(data = df_ell, aes(NMDS1, NMDS2, colour=mpa_type, group=group), size = 0.5, linetype=2 #aes(color=year)
            )+           #line connecting groups (years)
  #geom_point() +                                              #sample scores
  #coord_fixed() +
  #geom_text(data=cent, aes(label=year), position=position_jitter(width=0.02, height=0.01))+
  #xlim(-4.5,4.5)+
  #ylim(-0.45,0.65)+
  #facet_wrap(~mpa_designation,
  #           labeller = labeller(mpa_designation =
  #                                 c("ref"="REF",
  #                                   "smr"="SMR")))+
  ggtitle("kelp forest UPC")+
  scale_color_manual(values=c('red','blue'))+
  theme(strip.text = element_text(size=18, face="bold"))+
  theme(plot.title = element_text(size=16, face="bold"))+
  theme(text = element_text(20))
  )


```


#Kelp fish scores and plot
```{r}

scrs<- scores(kelp_fish_ord)
scrs <- cbind(as.data.frame(scrs), desig_state=kelp_fish_group_vars$desig_state, year=kelp_fish_group_vars$year, MHW=kelp_fish_group_vars$MHW, region4=kelp_fish_group_vars$region4, mpa_designation=kelp_fish_group_vars$mpa_defacto_designation) #to facilitate computing centroids, add group var

cent <- aggregate(cbind(NMDS1, NMDS2) ~ MHW+region4+mpa_designation, data = scrs, FUN = mean) #computes centroids by MHW and region




NMDS = data.frame(MDS1 = kelp_fish_ord$points[,1], MDS2 = kelp_fish_ord$points[,2],group=kelp_fish_group_vars$desig_state, mpa_desig=kelp_fish_group_vars$mpa_defacto_designation)


plot(kelp_fish_ord)
ord<- ordiellipse(kelp_fish_ord, kelp_fish_group_vars$desig_state, kind = "se",conf=0.95, label=T)

NMDS$group <- as.factor(NMDS$group)

df_ell <- data.frame()
for(g in levels(NMDS$group)){
  df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$group==g,],
                  vegan:::veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale)))
                                ,group=g))
}

df_ell$mpa_type <- word(df_ell$group, 1)

#ggplot(data = NMDS, aes(MDS1, MDS2)) + geom_point(aes(color = group)) +
#  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=group), size=1, linetype=2)


#reorder for plotting
cent$mpa_designation <- factor(cent$mpa_designation, level=c("smr","ref"))
cent$MHW <- factor(cent$MHW, level=c("before","during","after"))

(KF_fish <- ggplot(data=cent, aes(NMDS1, NMDS2)) +
  #geom_segment(data = segs,
               #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
  geom_point(aes(shape=MHW, color=mpa_designation), size=4) +                           #centroids
  geom_path(data = df_ell, aes(NMDS1, NMDS2, colour=mpa_type, group=group), size = 0.5, linetype=2 #aes(color=year)
            )+           #line connecting groups (years)
  #geom_point() +                                              #sample scores
  #coord_fixed() +
  #geom_text(data=cent, aes(label=year), position=position_jitter(width=0.02, height=0.01))+
  #xlim(-4.5,4.5)+
  #ylim(-0.45,0.65)+
  #facet_wrap(~mpa_designation,
  #           labeller = labeller(mpa_designation =
  #                                 c("ref"="REF",
  #                                   "smr"="SMR")))+
  ggtitle("kelp forest fish")+
  scale_color_manual(values=c('red','blue'))+
  theme(strip.text = element_text(size=18, face="bold"))+
  theme(plot.title = element_text(size=16, face="bold"))+
  theme(text = element_text(20))
)


```


#deep reef scores and plot
```{r}

scrs<- scores(deep_reef_ord)
scrs <- cbind(as.data.frame(scrs), desig_state=deep_reef_group_vars$desig_state, year=deep_reef_group_vars$year, MHW=deep_reef_group_vars$MHW, region4=deep_reef_group_vars$region4, mpa_designation=deep_reef_group_vars$mpa_defacto_designation) #to facilitate computing centroids, add group var

cent <- aggregate(cbind(NMDS1, NMDS2) ~ MHW+region4+mpa_designation, data = scrs, FUN = mean) #computes centroids by MHW and region




NMDS = data.frame(MDS1 = deep_reef_ord$points[,1], MDS2 = deep_reef_ord$points[,2],group=deep_reef_group_vars$desig_state, mpa_desig=deep_reef_group_vars$mpa_defacto_designation)


plot(deep_reef_ord)
ord<- ordiellipse(deep_reef_ord, deep_reef_group_vars$desig_state, kind = "se",conf=0.95, label=T)

NMDS$group <- as.factor(NMDS$group)

df_ell <- data.frame()
for(g in levels(NMDS$group)){
  df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$group==g,],
                  vegan:::veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale)))
                                ,group=g))
}

df_ell$mpa_type <- word(df_ell$group, 1)

#ggplot(data = NMDS, aes(MDS1, MDS2)) + geom_point(aes(color = group)) +
#  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=group), size=1, linetype=2)


#reorder for plotting
cent$mpa_designation <- factor(cent$mpa_designation, level=c("smr","ref"))
cent$MHW <- factor(cent$MHW, level=c("before","during","after"))

(deep_reef <- ggplot(data=cent, aes(NMDS1, NMDS2)) +
  #geom_segment(data = segs,
               #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
  geom_point(aes(shape=MHW, color=mpa_designation), size=4) +                           #centroids
  geom_path(data = df_ell, aes(NMDS1, NMDS2, colour=mpa_type, group=group), size = 0.5, linetype=2 #aes(color=year)
            )+           #line connecting groups (years)
  #geom_point() +                                              #sample scores
  #coord_fixed() +
  #geom_text(data=cent, aes(label=year), position=position_jitter(width=0.02, height=0.01))+
  #xlim(-4.5,4.5)+
  #ylim(-0.45,0.65)+
  #facet_wrap(~mpa_designation,
  #           labeller = labeller(mpa_designation =
  #                                 c("ref"="REF",
  #                                   "smr"="SMR")))+
  ggtitle("deep reef fish")+
  scale_color_manual(values=c('red','blue'))+
  theme(strip.text = element_text(size=18, face="bold"))+
  theme(plot.title = element_text(size=16, face="bold"))+
  theme(text = element_text(20))
)


```


#rocky scores and plot
```{r}

scrs<- scores(rocky_ord)
scrs <- cbind(as.data.frame(scrs), desig_state=rocky_group_vars$desig_state, year=rocky_group_vars$year, MHW=rocky_group_vars$MHW, region4=rocky_group_vars$region4, mpa_designation=rocky_group_vars$mpa_designation) #to facilitate computing centroids, add group var

cent <- aggregate(cbind(NMDS1, NMDS2) ~ MHW+region4+mpa_designation, data = scrs, FUN = mean) #computes centroids by MHW and region




NMDS = data.frame(MDS1 = rocky_ord$points[,1], MDS2 = rocky_ord$points[,2], group=rocky_group_vars$desig_state, mpa_desig=rocky_group_vars$mpa_designation)


plot(rocky_ord)
ord<- ordiellipse(rocky_ord, rocky_group_vars$desig_state, kind = "se",conf=0.95, label=T)

NMDS$group <- as.factor(NMDS$group)

df_ell <- data.frame()
for(g in levels(NMDS$group)){
  df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$group==g,],
                  vegan:::veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale)))
                                ,group=g))
}

df_ell$mpa_type <- word(df_ell$group, 1)

#ggplot(data = NMDS, aes(MDS1, MDS2)) + geom_point(aes(color = group)) +
#  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=group), size=1, linetype=2)


#reorder for plotting
cent$mpa_designation <- factor(cent$mpa_designation, level=c("smr","ref"))
cent$MHW <- factor(cent$MHW, level=c("before","during","after"))

(rocky <- ggplot(data=cent, aes(NMDS1, NMDS2)) +
  #geom_segment(data = segs,
               #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
  geom_point(aes(shape=MHW, color=mpa_designation), size=4) +                           #centroids
  geom_path(data = df_ell, aes(NMDS1, NMDS2, colour=mpa_type, group=group), size = 0.5, linetype=2 #aes(color=year)
            )+           #line connecting groups (years)
  #geom_point() +                                              #sample scores
  #coord_fixed() +
  #geom_text(data=cent, aes(label=year), position=position_jitter(width=0.02, height=0.01))+
  #xlim(-4.5,4.5)+
  #ylim(-0.45,0.65)+
  #facet_wrap(~mpa_designation,
  #           labeller = labeller(mpa_designation =
  #                                 c("ref"="REF",
  #                                   "smr"="SMR")))+
  ggtitle("rocky intertidal")+
  scale_color_manual(values=c('red','blue'))+
  theme(strip.text = element_text(size=18, face="bold"))+
  theme(plot.title = element_text(size=16, face="bold"))+
  theme(text = element_text(20))
)

```

#combine plots 

```{r}

figure<-ggarrange(CCFRP, KF_fish, deep_reef,
          KF_swath, KF_upc, rocky, common.legend = TRUE, legend="right",
          font.label=list(color="black",size=14)) + bgcolor('white')

annotate_figure(figure,
                top = text_grob("central CA community structure", color = "black", face = "bold", size = 20)
                #bottom = text_grob("Data source: \n ToothGrowth data set", color = "blue",
                                  # hjust = 1, x = 1, face = "italic", size = 10),
                #left = text_grob("Tooth length", color = "green", rot = 90),
                #right = "I'm done, thanks :-)!",
                #fig.lab = "Figure 1", fig.lab.face = "bold"
)

ggsave(here("analyses", "5community_climate_ecology", "figures", "community_structure_central.png"), figure, width = 12, units = "in", dpi = 300)
```



```{r}
#TEST CODE ORDIELLIPSE

CCFRP_process <- CCFRP_counts %>%
                  rowwise() %>%
                  dplyr::mutate(sum = sum(across(9:ncol(.)), na.rm = T)) %>%
                  filter(!(sum==0))%>%
                  dplyr::select(-sum)%>%
                  mutate(desig_state = paste(mpa_designation,MHW))%>%
                  dplyr::select(desig_state, everything())#remove rows containing only zeros
                  

#define grouping vars
CCFRP_group_vars <- CCFRP_process %>%
                    dplyr::select(1:9)

#define data for ordination
CCFRP_ord_data <- CCFRP_process %>%
              dplyr::select(10:ncol(.))


#calculate relative abundance
CCFRP_rel <- decostand(CCFRP_ord_data, method = "hellinger") %>%
            dplyr::select(where(~any(. !=0)))

#generate a BC dissim matrix
CCFRP_distmat <- vegdist(CCFRP_rel, method = "bray", na.rm=T) 




PCO <- capscale(CCFRP_distmat ~1)

plot(PCO, display="sites", type="n")
plot.window(xlim=c(-1,1), ylim=c(-1,1))

stats <- with(CCFRP_group_vars,
              ordiellipse(PCO,CCFRP_group_vars$desig_state,kind="se",conf=0.95, label=T,
                          col=c("red","blue","red","blue","red","blue")))

scrs<- scores(PCO, display="sites")
cent_MHW <- aggregate(scrs ~ desig_state, data=CCFRP_group_vars, FUN=mean)
names(cent_MHW)[-1] <- colnames(scrs)





points(cent_MHW[, -1], pch = 22, col='dark red', bg = c("red","blue","green"), cex = 1.1)

cent_year <- aggregate(scrs ~ year, data=CCFRP_group_vars, FUN=mean)
names(cent_year)[-1] <- colnames(scrs)

points(cent_year[, -1], pch = 22, col='blue', bg = c("red","blue","green"), cex = 1.1)






#https://stats.stackexchange.com/questions/34017/confidence-intervals-around-a-centroid-with-modified-gower-similarity

```
  



**Part 1.2 -- Calculate Centroid Distance**
#CCFRP_centroid distance
```{r}

CCFRP_process <- CCFRP_counts %>%
                  rowwise() %>%
                  dplyr::mutate(sum = sum(across(9:ncol(.)), na.rm = T)) %>%
                  filter(!(sum==0))%>%
                  dplyr::select(-sum) #remove rows containing only zeros

#define grouping vars
CCFRP_group_vars_smr <- CCFRP_process %>%
                    filter(mpa_designation=='smr')%>%
                    dplyr::select(1:8)

CCFRP_group_vars_ref <- CCFRP_process %>%
                    filter(mpa_designation=='ref')%>%
                    dplyr::select(1:8)

#define data for ordination
CCFRP_ord_smr <- CCFRP_process %>%
              filter(mpa_designation=='smr')%>%
              dplyr::select(9:ncol(.))

CCFRP_ord_ref <- CCFRP_process %>%
              filter(mpa_designation=='ref')%>%
              dplyr::select(9:ncol(.))

#calculate relative abundance
CCFRP_rel_smr <- decostand(CCFRP_ord_smr, method = "hellinger") %>%
            dplyr::select(where(~any(. !=0)))

CCFRP_rel_ref <- decostand(CCFRP_ord_ref, method = "hellinger") %>%
            dplyr::select(where(~any(. !=0)))


#generate a BC dissim matrix
CCFRP_distmat_smr <- vegdist(CCFRP_rel_smr, method = "bray", na.rm=T) 
CCFRP_distmat_ref <- vegdist(CCFRP_rel_ref, method = "bray", na.rm=T) 


(CCFRP_smr_before_after<- cbind("CCFRP",dist_between_centroids(CCFRP_distmat_smr, 1:27,39:54),"smr"))
(CCFRP_ref_before_after<- cbind("CCFRP",dist_between_centroids(CCFRP_distmat_ref, 1:27,39:54),"ref"))

(dist_traveled <- rbind(CCFRP_smr_before_after,CCFRP_ref_before_after))

```


#Kelp swath_centroid distance
```{r}

kelp_swath <- kelp_swath %>%
                  rowwise() %>%
                  dplyr::mutate(sum = sum(across(9:ncol(.)), na.rm = T)) %>%
                  filter(!(sum==0))%>%
                  dplyr::select(-sum) #remove rows containing only zeros

#define grouping vars
kelp_swath_group_vars_smr <- kelp_swath %>%
                    filter(mpa_defacto_designation=='smr')%>%
                    dplyr::select(1:8)

kelp_group_vars_ref <- kelp_swath %>%
                    filter(mpa_defacto_designation=='ref')%>%
                    dplyr::select(1:8)

#define data for ordination
kelp_ord_smr <- kelp_swath %>%
              filter(mpa_defacto_designation=='smr')%>%
              dplyr::select(9:ncol(.))

kelp_ord_ref <- kelp_swath %>%
              filter(mpa_defacto_designation=='ref')%>%
              dplyr::select(9:ncol(.))

#calculate relative abundance
kelp_rel_smr <- decostand(kelp_ord_smr, method = "hellinger") %>%
            dplyr::select(where(~any(. !=0)))

kelp_rel_ref <- decostand(kelp_ord_ref, method = "hellinger") %>%
            dplyr::select(where(~any(. !=0)))


#generate a BC dissim matrix
kelp_distmat_smr <- vegdist(kelp_rel_smr, method = "bray", na.rm=T) 
kelp_distmat_ref <- vegdist(kelp_rel_ref, method = "bray", na.rm=T) 


(kelp_smr_before_after<- cbind("kelp_swath",dist_between_centroids(kelp_distmat_smr, 1:101,110:124),"smr"))
(kelp_ref_before_after<- cbind("kelp_swath",dist_between_centroids(kelp_distmat_ref, 1:109,121:139),"ref"))

(dist_traveled <- rbind(dist_traveled,kelp_smr_before_after,kelp_ref_before_after))


```


Kelp upc_centroid distance
```{r}

kelp_upc <- kelp_upc_counts %>%
                  rowwise() %>%
                  dplyr::mutate(sum = sum(across(8:ncol(.), na.rm = T))) %>%
                  replace(is.na(.), 0)
                  

#define grouping vars
kelp_upc_vars_smr <- kelp_upc%>%
                    filter(mpa_defacto_designation=='smr')%>%
                    dplyr::select(1:7)
kelp_upc_vars_ref <- kelp_upc%>%
                    filter(mpa_defacto_designation=='ref')%>%
                    dplyr::select(1:7)

#define data for ordination
kelp_upc_ord_smr <- kelp_upc%>%
              filter(mpa_defacto_designation=='smr')%>%
              ungroup()%>%
              dplyr::select(8:ncol(.))

kelp_upc_ord_ref <- kelp_upc%>%
              filter(mpa_defacto_designation=='ref')%>%
              ungroup()%>%
              dplyr::select(8:ncol(.))

#calculate relative abundance
kelp_upc_rel_smr <- decostand(kelp_upc_ord_smr, method = "hellinger")
kelp_upc_rel_ref <- decostand(kelp_upc_ord_ref, method = "hellinger")

#generate a BC dissim matrix
kelp_upc_distmat_smr <- 
    vegdist(kelp_upc_rel_smr, method = "bray", na.rm=T) #generates a BC dissim matrix

kelp_upc_distmat_ref <- 
    vegdist(kelp_upc_rel_ref, method = "bray", na.rm=T)



(kelp_upc_smr_before_after<- cbind("kelp_upc",dist_between_centroids(kelp_upc_distmat_smr, 1:101,111:125),"smr"))
(kelp_upc_ref_before_after<- cbind("kelp_upc",dist_between_centroids(kelp_upc_distmat_ref, 1:109,121:139),"ref"))

(dist_traveled <- rbind(dist_traveled,kelp_upc_smr_before_after,kelp_upc_ref_before_after))


```


#kelp fish centroid distance
```{r}

kelp_fish <- kelp_fish_counts %>%
                  rowwise() %>%
                  dplyr::mutate(sum = sum(across(8:ncol(.), na.rm = T)))%>%
                  filter(!(sum==0))%>% #remove rows containing only zeros
                  dplyr::select(-sum, MHW, everything()) #drop sum columns and non-species categories

#define grouping vars
kelp_fish_vars_smr <- kelp_fish %>%
                    filter(mpa_defacto_designation=='smr')%>%
                    dplyr::select(1:7)

kelp_fish_vars_ref <- kelp_fish %>%
                    filter(mpa_defacto_designation=='ref')%>%
                    dplyr::select(1:7)

#define data for ordination
kelp_fish_ord_smr <- kelp_fish %>%
              filter(mpa_defacto_designation=='smr')%>%
              ungroup() %>%
              dplyr::select(8:ncol(.))

kelp_fish_ord_ref <- kelp_fish %>%
              filter(mpa_defacto_designation=='ref')%>%
              ungroup() %>%
              dplyr::select(8:ncol(.))


#calculate relative abundance
kelp_fish_rel_smr <- decostand(kelp_fish_ord_smr, method = "hellinger")
kelp_fish_rel_ref <- decostand(kelp_fish_ord_ref, method = "hellinger")


#generate a BC dissim matrix
kelp_fish_distmat_smr <- 
    vegdist(kelp_fish_rel_smr, method = "bray", na.rm=T) #generates a BC dissim matrix
kelp_fish_distmat_ref <- 
    vegdist(kelp_fish_rel_ref, method = "bray", na.rm=T) 


(kelp_fish_distmat_smr<- cbind("kelp_fish",dist_between_centroids(kelp_fish_distmat_smr, 1:100,110:124),"smr"))
(kelp_fish_distmat_ref <- cbind("kelp_fish",dist_between_centroids(kelp_upc_distmat_ref, 1:106,118:137),"ref"))

(dist_traveled <- rbind(dist_traveled,kelp_fish_distmat_smr,kelp_fish_distmat_ref))

```

#Deep reef centroid
```{r}
#deep reef processing---------------------------------------------------------

deep_reef <- deep_reef_counts %>%
                  rowwise() %>%
                  dplyr::mutate(sum = sum(across(8:ncol(.), na.rm = T))) %>%
                  filter(!(sum==0))%>% #remove rows containing only zeros
                  dplyr::select(-c(sum, schooling_10_15_cm_sebastes_sp,
                                schooling_10_15_cm_sebastes_sp,
                                young_of_year_10_cm_sebastes_sp,
                                synodus_lucioceps_or_ophiodon_elongatus,
                                sebastes_melanops_or_mystinus_or_diaconus)) #drop sum columns and non-species categories

#define grouping vars
deep_reef_vars_smr <- deep_reef%>%
                    filter(mpa_defacto_designation=='smr')%>%
                    dplyr::select(1:7)

deep_reef_vars_ref <- deep_reef%>%
                    filter(mpa_defacto_designation=='ref')%>%
                    dplyr::select(1:7)

#define data for ordination
deep_reef_ord_smr <- deep_reef%>%
              filter(mpa_defacto_designation=='smr')%>%
              ungroup()%>%
              dplyr::select(8:ncol(.))

deep_reef_ord_ref <- deep_reef%>%
              filter(mpa_defacto_designation=='ref')%>%
              ungroup()%>%
              dplyr::select(8:ncol(.))

#calculate relative abundance
deep_reef_rel_smr <- decostand(deep_reef_ord_smr, method="hellinger")
deep_reef_rel_ref <- decostand(deep_reef_ord_ref, method="hellinger")

#generate a BC dissim mat
deep_reef_distmat_smr <- 
    vegdist(deep_reef_rel_smr, method = "bray", na.rm=T) #generates a BC dissim matrix
deep_reef_distmat_ref <- 
    vegdist(deep_reef_rel_ref, method = "bray", na.rm=T) #generates a BC dissim matrix



(deep_reef_smr_before_after<- cbind("deep_reef",dist_between_centroids(deep_reef_distmat_smr, 1:15,16:22),"smr"))
(deep_reef_ref_before_after<- cbind("deep_reef",dist_between_centroids(deep_reef_distmat_ref, 1:7,16:25),"ref"))

(dist_traveled <- rbind(dist_traveled,deep_reef_smr_before_after,deep_reef_ref_before_after))

```




#intertidal 
```{r}

rocky_counts <- rocky_counts %>%
                mutate(MHW = ifelse(year>=2014 & year<=2016, "during",ifelse(year<2014, "before","after")))%>%
                dplyr::select(MHW, everything())%>%
                arrange(MHW)
                
                

#define grouping vars
rocky_group_vars_smr <- rocky_counts%>%
                    filter(mpa_designation=='smr')%>%
                    dplyr::select(1:9)

rocky_group_vars_ref <- rocky_counts%>%
                    filter(mpa_designation=='ref')%>%
                    dplyr::select(1:9)

#define data for ordination
rocky_ord_smr <- rocky_counts %>%
              filter(mpa_designation=='smr')%>%
              ungroup() %>%
              dplyr::select(10:ncol(.))

rocky_ord_ref <- rocky_counts %>%
              filter(mpa_designation=='ref')%>%
              ungroup() %>%
              dplyr::select(10:ncol(.))


#calculate relative abundance
rocky_rel_smr <- decostand(rocky_ord_smr, method = "hellinger")
rocky_rel_ref <- decostand(rocky_ord_ref, method = "hellinger")

#generate a BC dissim matrix
rocky_distmat_smr <- 
    vegdist(rocky_rel_smr, method = "bray", na.rm=T) #generates a BC dissim matrix

rocky_distmat_ref <- 
    vegdist(rocky_rel_ref, method = "bray", na.rm=T) 



(rocky_smr_before_after<- cbind("rocky", dist_between_centroids(rocky_distmat_smr, 1:21,22:101),"smr"))
(rocky_ref_before_after<- cbind("rocky", dist_between_centroids(rocky_distmat_ref, 1:33,34:144),"ref"))

(dist_traveled <- rbind(dist_traveled,rocky_smr_before_after,rocky_ref_before_after))


```

#create DF and plot
```{r}
dist_df = as.data.frame(dist_traveled)
colnames(dist_df)= c("group","distance","MPA") 

dist_df$community_type = c("fish","fish","inverts_algae","inverts_algae","inverts_algae","inverts_algae","fish","fish","fish","fish","inverts_algae","inverts_algae")


dist_df$group <- factor(dist_df$group, levels=c("kelp_swath","deep_reef","kelp_upc","kelp_fish","CCFRP","rocky"))
dist_df$MPA <- factor(dist_df$MPA, levels=c("ref","smr"))


A<- dist_df %>%
ggplot(aes(x=group, y=as.numeric(distance), fill=MPA))+
  geom_bar(stat='identity', color='black', position=position_dodge())+
  scale_fill_manual(values=c('#95B4CC','#ff392e'))+
  scale_y_continuous(labels=label_number(accuracy=0.01), 
                     breaks = seq(0,0.45,len=10))+
  coord_flip()+
  labs(
    title = "Distance traveled before/after MHW",
    x = "monitoring group",
    y = "distance (Euclidean)"
  )+
  #theme(legend.text = element_text(size=16))+
  theme_minimal(base_size=20)



dist_df_wide <- dist_df%>%
                pivot_wider(names_from = MPA, values_from='distance')%>%
                mutate(logRR = log10((as.numeric(smr)/as.numeric(ref))),
                       difference = as.numeric(smr)-as.numeric(ref),
                       group = reorder(group, logRR))
                
    


B <- dist_df_wide %>%
ggplot()+
  geom_bar(mapping = aes(x=group, y=logRR, fill=community_type), stat='identity', color='black', position=position_dodge())+
  #scale_fill_manual(values=c('#95B4CC','#ff392e'))+
  #scale_y_continuous(labels=label_number(accuracy=0.01), 
  #                   breaks = seq(0,0.45,len=10))+
  #xlim(-0.2,0.10)+
  scale_y_continuous(limits=c(-0.20,0.10))+
  coord_flip()+
  scale_fill_brewer(palette = "Set2")+
  labs(
    title = "Distance traveled response ratio of distance traveled",
    x = "monitoring group",
    y = "response ratio (log(smr/ref))"
  )+
  theme_minimal(base_size=20)


ggarrange(A,B,
          labels="AUTO",
          font.label = list(size=16))


```










**PART 1.3 COMMUNITY TRAJECTORY OVER TIME**


#CCFRP scores and plot
```{r}


scrs <- scores(CCFRP_ord)

scrs <- cbind(as.data.frame(scrs), desig_state=CCFRP_group_vars$desig_state, year=CCFRP_group_vars$year, MHW=CCFRP_group_vars$MHW, region4=CCFRP_group_vars$region4, mpa_designation=CCFRP_group_vars$mpa_designation) #to facilitate computing centroids, add group var

scrs$region4 <- factor(scrs$region4, level = c("north","central","north islands","south")) #set order
scrs$MHW <- factor(scrs$MHW, level = c("before","during","after")) #set order
scrs$mpa_designation <- factor(scrs$mpa_designation, level = c("smr","ref")) #set order


cent <- aggregate(cbind(NMDS1, NMDS2) ~ year+MHW+region4+mpa_designation, data = scrs, FUN = mean) #computes centroids by MHW and region




CCFRP_trajectory<- ggplot(data=cent, aes(NMDS1, NMDS2)) +
  #geom_segment(data = segs,
               #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
  geom_point(aes(shape=MHW, color=mpa_designation), size=4) +                           #centroids
  geom_path(data=cent, aes(NMDS1, NMDS2, group=mpa_designation, color=mpa_designation), size = 0.5, linetype=1 #aes(color=year)
            )+           #line connecting groups (years)
  ggtitle("CCFRP")+
  scale_color_manual(values=c('red','blue'))+
  theme(strip.text = element_text(size=18, face="bold"))+
  theme(plot.title = element_text(size=16, face="bold"))+
  theme(text = element_text(20))
  #geom_text(position=position_jitter(width=0.01, height=0.05))


CCFRP_trajectory

```



#kelp_swath scores and plot
```{r}

scrs <- scores(kelp_swath_ord)

scrs <- cbind(as.data.frame(scrs), desig_state= kelp_swath_group_vars$desig_state, year=kelp_swath_group_vars$year, MHW=kelp_swath_group_vars$MHW, region4=kelp_swath_group_vars$region4, mpa_designation=kelp_swath_group_vars$mpa_defacto_designation) #to facilitate computing centroids, add group var

scrs$region4 <- factor(scrs$region4, level = c("north","central","north islands","south")) #set order
scrs$MHW <- factor(scrs$MHW, level = c("before","during","after")) #set order
scrs$mpa_designation <- factor(scrs$mpa_designation, level = c("smr","ref")) #set order


cent <- aggregate(cbind(NMDS1, NMDS2) ~ year+MHW+region4+mpa_designation, data = scrs, FUN = mean) #computes centroids by MHW and region

kelp_swath_trajectory<- ggplot(data=cent, aes(NMDS1, NMDS2, label=year)) +
  #geom_segment(data = segs,
               #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
  geom_point(aes(shape=MHW, color=mpa_designation), size=4) +                           #centroids
  geom_path(data=cent, aes(NMDS1, NMDS2, group=mpa_designation, color=mpa_designation), size = 0.5, linetype=1 #aes(color=year)
            )+           #line connecting groups (years)
 
  ggtitle("kelp swath")+
  scale_color_manual(values=c('red','blue'))+
  theme(strip.text = element_text(size=18, face="bold"))+
  theme(plot.title = element_text(size=16, face="bold"))+
  theme(text = element_text(20))
  #geom_text(position=position_jitter(width=0.01, height=0.05))


kelp_swath_trajectory


```



#kelp_upc scores and plot
```{r}

scrs <- scores(kelp_upc_ord)

scrs <- cbind(as.data.frame(scrs), desig_state= kelp_upc_group_vars$desig_state, year=kelp_upc_group_vars$year, MHW=kelp_upc_group_vars$MHW, region4=kelp_upc_group_vars$region4, mpa_designation=kelp_upc_group_vars$mpa_defacto_designation) #to facilitate computing centroids, add group var

scrs$region4 <- factor(scrs$region4, level = c("north","central","north islands","south")) #set order
scrs$MHW <- factor(scrs$MHW, level = c("before","during","after")) #set order
scrs$mpa_designation <- factor(scrs$mpa_designation, level = c("smr","ref")) #set order


cent <- aggregate(cbind(NMDS1, NMDS2) ~ year+MHW+region4+mpa_designation, data = scrs, FUN = mean) #computes centroids by MHW and region

kelp_upc_trajectory<- ggplot(data=cent, aes(NMDS1, NMDS2, label=year)) +
  #geom_segment(data = segs,
               #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
  geom_point(aes(shape=MHW, color=mpa_designation), size=4) +                           #centroids
  geom_path(data=cent, aes(NMDS1, NMDS2, group=mpa_designation, color=mpa_designation), size = 0.5, linetype=1 #aes(color=year)
            )+           #line connecting groups (years)
 
  ggtitle("kelp UPC")+
  scale_color_manual(values=c('red','blue'))+
  theme(strip.text = element_text(size=18, face="bold"))+
  theme(plot.title = element_text(size=16, face="bold"))+
  theme(text = element_text(20))
  #geom_text(position=position_jitter(width=0.01, height=0.05))


kelp_upc_trajectory


```


#kelp_fish scores and plot
```{r}

scrs <- scores(kelp_fish_ord)

scrs <- cbind(as.data.frame(scrs), desig_state= kelp_fish_group_vars$desig_state, year=kelp_fish_group_vars$year, MHW=kelp_fish_group_vars$MHW, region4=kelp_fish_group_vars$region4, mpa_designation=kelp_fish_group_vars$mpa_defacto_designation) #to facilitate computing centroids, add group var

scrs$region4 <- factor(scrs$region4, level = c("north","central","north islands","south")) #set order
scrs$MHW <- factor(scrs$MHW, level = c("before","during","after")) #set order
scrs$mpa_designation <- factor(scrs$mpa_designation, level = c("smr","ref")) #set order


cent <- aggregate(cbind(NMDS1, NMDS2) ~ year+MHW+region4+mpa_designation, data = scrs, FUN = mean) #computes centroids by MHW and region

kelp_fish_trajectory<- ggplot(data=cent, aes(NMDS1, NMDS2, label=year)) +
  #geom_segment(data = segs,
               #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
  geom_point(aes(shape=MHW, color=mpa_designation), size=4) +                           #centroids
  geom_path(data=cent, aes(NMDS1, NMDS2, group=mpa_designation, color=mpa_designation), size = 0.5, linetype=1 #aes(color=year)
            )+           #line connecting groups (years)
 
  ggtitle("kelp UPC")+
  scale_color_manual(values=c('red','blue'))+
  theme(strip.text = element_text(size=18, face="bold"))+
  theme(plot.title = element_text(size=16, face="bold"))+
  theme(text = element_text(20))
  #geom_text(position=position_jitter(width=0.01, height=0.05))


kelp_fish_trajectory


```



#deep_reef scores and plot
```{r}

scrs <- scores(deep_reef_ord)

scrs <- cbind(as.data.frame(scrs), desig_state= deep_reef_group_vars$desig_state, year=deep_reef_group_vars$year, MHW=deep_reef_group_vars$MHW, region4=deep_reef_group_vars$region4, mpa_designation=deep_reef_group_vars$mpa_defacto_designation) #to facilitate computing centroids, add group var

scrs$region4 <- factor(scrs$region4, level = c("north","central","north islands","south")) #set order
scrs$MHW <- factor(scrs$MHW, level = c("before","during","after")) #set order
scrs$mpa_designation <- factor(scrs$mpa_designation, level = c("smr","ref")) #set order


cent <- aggregate(cbind(NMDS1, NMDS2) ~ year+MHW+region4+mpa_designation, data = scrs, FUN = mean) #computes centroids by MHW and region

deep_reef_trajectory<- ggplot(data=cent, aes(NMDS1, NMDS2, label=year)) +
  #geom_segment(data = segs,
               #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
  geom_point(aes(shape=MHW, color=mpa_designation), size=4) +                           #centroids
  geom_path(data=cent, aes(NMDS1, NMDS2, group=mpa_designation, color=mpa_designation), size = 0.5, linetype=1 #aes(color=year)
            )+           #line connecting groups (years)
 
  ggtitle("deep reef")+
  scale_color_manual(values=c('red','blue'))+
  theme(strip.text = element_text(size=18, face="bold"))+
  theme(plot.title = element_text(size=16, face="bold"))+
  theme(text = element_text(20))
  #geom_text(position=position_jitter(width=0.01, height=0.05))


deep_reef_trajectory


```


#rocky scores and plot
```{r}

scrs <- scores(rocky_ord)

scrs <- cbind(as.data.frame(scrs), desig_state= rocky_group_vars$desig_state, year=rocky_group_vars$year, MHW=rocky_group_vars$MHW, region4=rocky_group_vars$region4, mpa_designation=rocky_group_vars$mpa_designation) #to facilitate computing centroids, add group var

scrs$region4 <- factor(scrs$region4, level = c("north","central","north islands","south")) #set order
scrs$MHW <- factor(scrs$MHW, level = c("before","during","after")) #set order
scrs$mpa_designation <- factor(scrs$mpa_designation, level = c("smr","ref")) #set order


cent <- aggregate(cbind(NMDS1, NMDS2) ~ year+MHW+region4+mpa_designation, data = scrs, FUN = mean) #computes centroids by MHW and region

rocky_trajectory<- ggplot(data=cent, aes(NMDS1, NMDS2, label=year)) +
  #geom_segment(data = segs,
               #mapping = aes(xend = oNMDS1, yend = oNMDS2)) + #spiders
  geom_point(aes(shape=MHW, color=mpa_designation), size=4) +                           #centroids
  geom_path(data=cent, aes(NMDS1, NMDS2, group=mpa_designation, color=mpa_designation), size = 0.5, linetype=1 #aes(color=year)
            )+           #line connecting groups (years)
 
  ggtitle("rocky intertidal")+
  scale_color_manual(values=c('red','blue'))+
  theme(strip.text = element_text(size=18, face="bold"))+
  theme(plot.title = element_text(size=16, face="bold"))+
  theme(text = element_text(20))
  #geom_text(position=position_jitter(width=0.01, height=0.05))


rocky_trajectory


```


#combine plots 
```{r}
require('ggpubr')
figure<-ggarrange(CCFRP_trajectory, kelp_fish_trajectory, deep_reef_trajectory,
          kelp_swath_trajectory, kelp_upc_trajectory, rocky_trajectory, common.legend = TRUE, legend="right",
          font.label=list(size=40)) + bgcolor('white')
annotate_figure(figure,
                top = text_grob("central CA community structure", color = "black", face = "bold", size = 20)
)

#Export figure
ggsave(here("analyses", "5community_climate_ecology", "figures", "community_trajectory.png"), figure, width = 12, units = "in", dpi = 300)


#png("/home/joshsmith/CA_MPA_Project/ca-mpa/analyses/5community_climate_ecology/figures/community_trajectory.png")
#aurora = "/home/joshsmith/CA_MPA_Project/ca-mpa/analyses/5community_climate_ecology/figures"
#ggsave(plot=last_plot(), path = aurora, device='png',filename = "community_trajectory.png", scale=1, dpi=72)
#dev.off()
```



