---
title: "PCA for MPA Attribute Data"
author: "Cori Lopazanski"
date: '2022-06-08'
output: 
  html_document:
    self_contained: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```


# Overview

*Goal*: Explore whether key characteristics describe variation among MPAs. Can
we reduce the attribute data to fewer variables while still capturing the
key information? Are there useful groupings that emerge which might be useful
for subsequent analyses? Are there different drivers of variation in different regions?

*Steps*
1. Log-transform the continuous data: standardize across variables with different scales
2. PCA with all variables 
3. Reduce variables which explain the same underlying characteristics
4. PCA with consolidated variables 
5. Separate regional PCAs with all/consolidated variables

*Data*

The MPA Attributes data for these analyses was supplied by California Department of Fish and Wildlife.
It is contained in the CA MPA Working Group shared data folder. 

Link: https://drive.google.com/drive/folders/1tCOIsQ2TolxAqZW13gdLjD_tEn27T9ox?usp=sharing


```{r}
# Packages
library(tidyverse)
library(here)
library(ggbiplot)
library(GGally)
library(vegan)
library(viridis)
library(RColorBrewer)
# Directories
base.dir <- "/Volumes/GoogleDrive/Shared drives/NCEAS MPA network assessment/MPA Network Assessment: Working Group Shared Folder/data/sync-data" # Cori Local
#base.dir <- "/home/shares/ca-mpa/data/sync-data/" #Aurora Base
repo.dir <- here::here("analyses", "7habitat")
in.dir <- here::here(base.dir, "mpa_traits", "processed")
fig.dir <- here::here("analyses", "7habitat", "figures")

# Read Cleaned Attribute Data
attributes <- read_csv(file.path(in.dir, "mpa_attributes_clean.csv"))

# Some Minor Processing
data <- attributes %>% 
  mutate(four_region_north_ci = as.factor(four_region_north_ci)) %>% 
  mutate(total_submarine_canyon_km2 = 
           rowSums(select(., c(submarine_canyon_0_30m_km2,
                               submarine_canyon_30_100m_km2,
                               submarine_canyon_100_200m_km2)))) %>%
  mutate(total_estuary = 
           rowSums(select(., c(estuary_km2, coastal_marsh_km2, eelgrass_km2,
                               coastal_marsh_km, tidal_flats_km, surfgrass_km)))) %>% 
  filter(mpa_class != "smrma")

```
# Summarize Biomass Data

## Log Response Ratios

```{r}
# Read Cleaned Biomass Response Ratio Data
biomass_logRR <- read_csv(file.path(base.dir, "monitoring", "processed_data", 
                              "targeted_nontargeted_fish_biomass_logRRs.csv"))

# Average response for each group-site for 2016-2019
biomass_1619_logRR <- biomass_logRR %>% 
  filter(year %in% c(2016:2019)) %>% 
  group_by(group, affiliated_mpa, target_status) %>% 
  dplyr::summarize(mean_logRR = mean(logRR, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(group_ind = paste(group, target_status, "logRR", sep = "_")) %>% 
  select(!c(group, target_status)) %>% 
  pivot_wider(names_from = group_ind, values_from = mean_logRR)

```

## True Biomass Values (not calculated response ratios)

```{r}
biomass <- read_csv("/Volumes/GoogleDrive/Shared drives/NCEAS MPA network assessment/MPA Network Assessment: Working Group Shared Folder/data/sync-data/monitoring/processed_data/targeted_nontargeted_fish_biomass.csv")

# Average biomass for each group-site for 2016-2019
# MPAs only

biomass_1619 <- biomass %>% 
  filter(mpa_designation != "ref",
         year %in% c(2016:2019)) %>% 
  group_by(group, affiliated_mpa, target_status) %>% 
  dplyr::summarize(mean_total_annual_biomass = mean(sum_biomass, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(group_ind = paste(group, target_status, "total", sep = "_")) %>% 
  select(!c(group, target_status)) %>% 
  pivot_wider(names_from = group_ind, values_from = mean_total_annual_biomass)

```


# Log-Transform Attribute Data

Otherwise, variables with high variances (e.g. depth range) will have higher loadings in the principal components. Since the data is zero-heavy, I am using a log+1 transformation to all continuous variables.

```{r}
log_data1 <- data %>% 
  select(name, 
         mpa_class, 
         four_region_north_ci,
         size_km2, distance_to_port, shore_span_km,
         sandy_beach_km:tidal_flats_km, 
         depth_range, 
         total_hard_substrate, 
         total_soft_substrate,
         total_submarine_canyon_km2,
         total_estuary) %>% 
  select(!max_kelp_canopy_landsat_km2) %>% # Drop because too many NAs 
  drop_na() %>% # Remove NA values because na.omit not working in princomp()
  mutate_if(., is.numeric, log1p)

```


```{r}
# Pairwise correlations
# Warning: this is hectic and takes a minute
#ggpairs(log_data[4:35], aes(col = log_data$four_region_north_ci, alpha = 0.8))

# Some of the data is still very skewed due to high number of zeroes
# Some also has pretty high correlation - e.g. depth
```

# Join Attribute and Biomass Data



```{r}
# Join with logRR
log_data2 <- left_join(log_data1, biomass_1619_logRR, by = c("name" = "affiliated_mpa")) 

# Join with mean biomass values
log_data3 <- left_join(log_data2, biomass_1619, by = c("name" = "affiliated_mpa"))

# Rename for phew clarity
log_data <- log_data3 

```


# PCA: All Continuous Variables

First analysis includes all continuous variables related to the underlying habitat traits:

```{r}
names(log_data[4:31])
```




```{r all-vars, fig.width = 8, fig.height = 6}
# Generate principal components
pca <- prcomp(log_data[4:31],
              center = TRUE,
              scale = TRUE) # normalize with standard deviation 1

# Plot
ggbiplot(pca, group = log_data$four_region_north_ci, ellipse = T,
         obs.scale = 1, var.scale = 1) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Region")

# Examine loadings
loadings <- as.data.frame(pca$rotation)

```
**Fig 1.** Principal component analysis biplot for all MPA attribute variables linked to habitat characteristics. 


## Explore Explained Variance
```{r}
# Retrieve the standard deviation of each PC
stdev <- pca$sdev

# Calculate variance of each PC
var <- stdev^2

# Calculate proportion of variance explained by each PC
prop_var_ex <- var/sum(var) 
prop_var_ex[1:20]

# Scree plot shows we get past 98% var explained ~ PC 25
plot(cumsum(prop_var_ex), 
     xlab = "PC", 
     ylab = "Cumulative proportion of variance explained", 
     type = "b")
abline(h = 0.98)

# Can also get this through summary
summary(pca)
```

## Reducing Variables

PCA estimates the contribution of different variables to the variation among the points - if there is high correlation between some of the variables, that would result in a higher weight of those variables in the estimation of the components. If the correlated variables are each describing different aspects of the data, then it would be wise to retain them. However, if there are multiple variables that are measuring the same underlying aspect of the data, it could be worthwhile to discard or consolidate them.

In the attribute data, there are several variables for the amount of hard/soft substrate calculated in different depth ranges. This could be useful in describing the spatial allocation of habitat in different depth zones, but could also be redundant if the attribute of interest is the general amount of hard or soft bottom, given that there is another variable which describes the depth range of the entire area. Similarly, there are several variables which describe estuarine characteristics.

(Potentially) Reasonable Decisions for Consolidating Variables: 

- Consolidate minimum and maximum depth to a single “depth range” variable
- Consolidate the hard substrate columns to a single “total hard substrate” variable
- Consolidate the soft bottom columns to a single “total soft substrate” variable
- Consolidate the submarine canyon columns to a single “total submarine canyon” variable
- Remove duplicates among estuarine habitat types and/or calculate a “total estuarine” variable


```{r}
log_data_reduced <- log_data %>% 
  select(name, 
         mpa_class, 
         four_region_north_ci, 
         distance_to_port, 
         size_km2, 
         shore_span_km, 
         sandy_beach_km, 
         rocky_inter_km, 
         offshore_rock_km, 
         max_kelp_canopy_cdfw_km2,
         depth_range,
         total_hard_substrate,
         total_soft_substrate,
         total_submarine_canyon_km2,
         total_estuary)
```


## Reduced PCA

Variables:

```{r}
names(log_data_reduced[4:15])
```

```{r reduced-vars, fig.width = 8, fig.height = 6}
pca_reduced <- prcomp(log_data_reduced[4:15],
                      center = TRUE,
                      scale = TRUE) 

summary(pca_reduced)

plot(pca_reduced, type = "lines")

ggbiplot(pca_reduced, 
         group = log_data_reduced$four_region_north_ci, 
         ellipse = T, 
         obs.scale = 1, 
         var.scale = 1,) +
  theme_minimal() +
  labs(title = "Reduced Trait Variables")

# Principal component loadings
loadings_reduced <- as.data.frame(pca_reduced$rotation)

```


```{r}
# This is hacky and I'm not sure if it's recommended
# See which variables contribute the most...
#loadings_reduced$total <- rowSums(abs(loadings_reduced[2:13]))
#loadings_reduced$total_first10 <- rowSums(abs(loadings_reduced[2:11]))

#loadings$total <- rowSums(abs(loadings[2:39]))
#loadings$total_first10 <- rowSums(abs(loadings[2:11]))
```

# PCA: Regional

There could be regional differences in the primary drivers of variation for the MPAs. Do the PCA again, except each region separately (North, Central, South, Northern Islands)


## North

```{r north, fig.width = 8, fig.height= 6}
north   <- log_data %>% 
  filter(four_region_north_ci == "north") %>% 
  select(!submarine_canyon_0_30m_km2) # all zeroes

pca_n <-  prcomp(north[4:30], center = TRUE, scale = TRUE)

ggbiplot(pca_n, labels = north$name,
         obs.scale = 1, var.scale = 1,
         group = north$mpa_class) + 
  scale_x_continuous(limits = c(-10, 3)) +
  scale_y_continuous(limits = c(-10, 8))+
  labs(title = "North - All Variables") +
  theme_minimal()

ggbiplot(pca_n, labels = north$name,
         obs.scale = 1, var.scale = 1,
         group = north$mpa_class,
         var.axes = FALSE) + 
  scale_x_continuous(limits = c(-10, 9)) +
  scale_y_continuous(limits = c(-10, 2))+
  labs(title = "North - All Variables") +
  theme_minimal()

north_reduced <- log_data_reduced %>% 
  filter(four_region_north_ci == "north")

pca_n2 <-  prcomp(north_reduced[4:15], center = TRUE, scale = TRUE)

ggbiplot(pca_n2, 
         labels = north_reduced$name,
         obs.scale = 1, var.scale = 1,
         group = north_reduced$mpa_class) +
  scale_x_continuous(limits = c(-4, 4)) +
  scale_y_continuous(limits = c(-4, 4))+
  labs(title = "North - Reduced Variables") +
  theme_minimal()

ggbiplot(pca_n2, 
         labels = north_reduced$name,
         obs.scale = 1, var.scale = 1,
         group = north_reduced$mpa_class,
         var.axes = FALSE) +
  scale_x_continuous(limits = c(-4, 4)) +
  scale_y_continuous(limits = c(-3.5, 3))+
  labs(title = "North - Reduced Variables") +
  theme_minimal()
```

## Central

```{r central, fig.width = 8, fig.height = 6}
central <- log_data %>% 
  filter(four_region_north_ci == "central")

pca_c <-  prcomp(central[4:35], center = TRUE, scale = TRUE)

ggbiplot(pca_c, labels = central$name,
         obs.scale = 1, var.scale = 1,
         group = central$mpa_class) +
  scale_x_continuous(limits = c(-8.5, 8.5)) +
  scale_y_continuous(limits = c(-6, 9))+
  labs(title = "Central - All Variables") +
  theme_minimal()

ggbiplot(pca_c, labels = central$name,
         obs.scale = 1, var.scale = 1,
         group = central$mpa_class, var.axes = FALSE) +
  scale_x_continuous(limits = c(-6, 6)) +
  scale_y_continuous(limits = c(-4, 6))+
  labs(title = "Central - All Variables") +
  theme_minimal()

central_reduced <- log_data_reduced %>% 
  filter(four_region_north_ci == "central")

pca_c2 <-  prcomp(central_reduced[4:15], center = TRUE, scale = TRUE)

ggbiplot(pca_c2, labels = central_reduced$name,
         obs.scale = 1, var.scale = 1, 
         group = central_reduced$mpa_class) +
    labs(title = "Central - Reduced Variables") +
  scale_x_continuous(limits = c(-5, 5)) +
  scale_y_continuous(limits = c(-5, 4))+
    theme_minimal()

ggbiplot(pca_c2, labels = central_reduced$name,
         obs.scale = 1, var.scale = 1, 
         group = central_reduced$mpa_class,
         var.axes = F) +
    labs(title = "Central - Reduced Variables") +
  scale_x_continuous(limits = c(-5, 5)) +
  scale_y_continuous(limits = c(-4, 2.5))+
    theme_minimal()
```

## South
```{r south, fig.width = 8, fig.height = 6}
south <- log_data %>% 
  filter(four_region_north_ci == "south") %>% 
  select(!eelgrass_km2) # all zeroes

pca_s <-  prcomp(south[4:34], center = TRUE, scale = TRUE)

ggbiplot(pca_s, labels = south$name,
         obs.scale = 1, var.scale = 1, group = south$mpa_class) +
  scale_y_continuous(limits = c(-9, 3))+
  scale_x_continuous(limits = c(-5, 8))+
    labs(title = "South - All Variables") +
    theme_minimal()

ggbiplot(pca_s, labels = south$name,
         obs.scale = 1, var.scale = 1, 
         group = south$mpa_class,
         var.axes = F) +
  scale_y_continuous(limits = c(-8.5, 2.5))+
  scale_x_continuous(limits = c(-5, 6))+
    labs(title = "South - All Variables") +
    theme_minimal()

south_reduced <- log_data_reduced %>% 
  filter(four_region_north_ci == "south")

pca_s2 <-  prcomp(south_reduced[4:15], center = TRUE, scale = TRUE)

ggbiplot(pca_s2, labels = south_reduced$name,
         obs.scale = 1, var.scale = 1, 
         group = south_reduced$mpa_class) +
  scale_y_continuous(limits = c(-3, 4))+
  scale_x_continuous(limits = c(-5, 6))+
    labs(title = "South - Reduced Variables") +
    theme_minimal()

ggbiplot(pca_s2, labels = south_reduced$name,
         obs.scale = 1, var.scale = 1,
         var.axes = F,
         group = south_reduced$mpa_class) +
  scale_x_continuous(limits = c(-5, 6.5))+
  scale_y_continuous(limits = c(-3, 2.5))+
    labs(title = "South - Reduced Variables") +
    theme_minimal()
```

## Islands

```{r n-islands, fig.width = 8, fig.height = 6}
islands <- log_data %>% 
  filter(four_region_north_ci == "north islands") %>% 
  select(!c(submarine_canyon_0_30m_km2:submarine_canyon_100_200m_km2,
            estuary_km2, coastal_marsh_km, tidal_flats_km)) # All zeroes

pca_i <-  prcomp(islands[4:29], center = TRUE, scale = TRUE)

ggbiplot(pca_i, labels = islands$name,
         obs.scale = 1, var.scale = 1,
         group = islands$mpa_class) +
  scale_x_continuous(limits = c(-6, 6))+
  scale_y_continuous(limits = c(-7.5, 5))+
  labs(title = "Northern Channel Islands - All Variables") +
  theme_minimal()

ggbiplot(pca_i, labels = islands$name,
         obs.scale = 1, var.scale = 1,
         group = islands$mpa_class, var.axes = F) +
  scale_x_continuous(limits = c(-6, 6))+
  scale_y_continuous(limits = c(-7.5, 5))+
  labs(title = "Northern Channel Islands - All Variables") +
  theme_minimal()
         
         

islands_reduced <- log_data_reduced %>% 
  filter(four_region_north_ci == "north islands")

pca_i2 <-  prcomp(islands_reduced[4:15], center = TRUE, scale = TRUE)

ggbiplot(pca_i2, labels = islands_reduced$name,
         obs.scale = 1, var.scale = 1,
         group = islands_reduced$mpa_class)+
  scale_x_continuous(limits = c(-4, 4))+
  scale_y_continuous(limits = c(-3, 4))+
    labs(title = "Northern Islands - Reduced Variables") +
    theme_minimal()   

ggbiplot(pca_i2, labels = islands_reduced$name,
         obs.scale = 1, var.scale = 1,
         group = islands_reduced$mpa_class,
         var.axes = F)+
  scale_x_continuous(limits = c(-4, 4))+
  scale_y_continuous(limits = c(-3, 4))+
    labs(title = "Northern Islands - Reduced Variables") +
    theme_minimal()  
```

# Incorporate Response Variables into PCA

## LogRR Targeted Fish Biomass

### Kelp

```{r kelp-targeted-logRR}
ggbiplot(pca, group = log_data$kelp_targeted_logRR,
         obs.scale = 1, var.scale = 1) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Kelp Targeted \nFish Biomass \nLogRR")


ggbiplot(pca, group = log_data$kelp_targeted_logRR,
         obs.scale = 1, var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Kelp Targeted \nFish Biomass \nLogRR")

ggbiplot(pca, group = log_data$kelp_targeted_logRR,
         obs.scale = 1, var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Kelp Targeted \nFish Biomass \nLogRR")

ggbiplot(pca_reduced, 
         group = log_data$kelp_targeted_logRR,
         obs.scale = 1, 
         var.scale = 1,) +
  theme_minimal() +
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "Reduced Trait Variables")

ggbiplot(pca_reduced, 
         group = log_data$kelp_targeted_logRR,
         obs.scale = 1, 
         var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "Reduced Trait Variables")

ggbiplot(pca_reduced, 
         group = log_data$kelp_targeted_logRR,
         obs.scale = 1, 
         var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "Reduced Trait Variables")

```

### CCFRP Targeted

```{r ccfrp-targeted-logRR}
ggbiplot(pca, group = log_data$ccfrp_targeted_logRR,
         obs.scale = 1, var.scale = 1) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "CCFRP Targeted \nFish Biomass \nLogRR")

ggbiplot(pca, group = log_data$ccfrp_targeted_logRR,
         obs.scale = 1, var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "CCFRP Targeted \nFish Biomass \nLogRR")

ggbiplot(pca, group = log_data$ccfrp_targeted_logRR,
         obs.scale = 1, var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "CCFRP Targeted \nFish Biomass \nLogRR")

ggbiplot(pca_reduced, 
         group = log_data$ccfrp_targeted_logRR,
         obs.scale = 1, 
         var.scale = 1,) +
  theme_minimal() +
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "CCFRP Targeted \nFish Biomass \nLogRR")

ggbiplot(pca_reduced, 
         group = log_data$ccfrp_targeted_logRR,
         obs.scale = 1, 
         var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "CCFRP Targeted \nFish Biomass \nLogRR")

ggbiplot(pca_reduced, 
         group = log_data$ccfrp_targeted_logRR,
         obs.scale = 1, 
         var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "CCFRP Targeted \nFish Biomass \nLogRR")
```

### Deep Reef Targeted

```{r deep-reef-targeted-logRR}
ggbiplot(pca, group = log_data$deep_reef_targeted_logRR,
         obs.scale = 1, var.scale = 1) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Deep Reef Targeted \nFish Biomass \nLogRR")

ggbiplot(pca, group = log_data$deep_reef_targeted_logRR,
         obs.scale = 1, var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Deep Reef Targeted \nFish Biomass \nLogRR")

ggbiplot(pca, group = log_data$deep_reef_targeted_logRR,
         obs.scale = 1, var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Deep Reef Targeted \nFish Biomass \nLogRR")

ggbiplot(pca_reduced, 
         group = log_data$deep_reef_targeted_logRR,
         obs.scale = 1, 
         var.scale = 1,) +
  theme_minimal() +
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "Deep Reef Targeted \nFish Biomass \nLogRR")

ggbiplot(pca_reduced, 
         group = log_data$deep_reef_targeted_logRR,
         obs.scale = 1, 
         var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "Deep Reef Targeted \nFish Biomass \nLogRR")

ggbiplot(pca_reduced, 
         group = log_data$deep_reef_targeted_logRR,
         obs.scale = 1, 
         var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "Deep Reef Targeted \nFish Biomass \nLogRR")
```
### Surf Zone Fish

```{r surf-targeted-logRR}
ggbiplot(pca, group = log_data$surf_targeted_logRR,
         obs.scale = 1, var.scale = 1) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Surf Targeted \nFish Biomass \nLogRR")

ggbiplot(pca, group = log_data$surf_targeted_logRR,
         obs.scale = 1, var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Surf Targeted \nFish Biomass \nLogRR")

ggbiplot(pca, group = log_data$surf_targeted_logRR,
         obs.scale = 1, var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Surf Targeted \nFish Biomass \nLogRR")

ggbiplot(pca_reduced, 
         group = log_data$surf_targeted_logRR,
         obs.scale = 1, 
         var.scale = 1,) +
  theme_minimal() +
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "Surf Targeted \nFish Biomass \nLogRR")

ggbiplot(pca_reduced, 
         group = log_data$surf_targeted_logRR,
         obs.scale = 1, 
         var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "Surf Targeted \nFish Biomass \nLogRR")

ggbiplot(pca_reduced, 
         group = log_data$surf_targeted_logRR,
         obs.scale = 1, 
         var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_colour_distiller(palette = "PuOr", na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "Surf Targeted \nFish Biomass \nLogRR")
```

## Actual Total Values Targeted Fish Biomass

### Kelp Forest

```{r kelp-targeted-total}
ggbiplot(pca, group = log_data$kelp_targeted_total,
         obs.scale = 1, var.scale = 1) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_continuous(na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Kelp Targeted \nFish Biomass \nMean Total Annual")


ggbiplot(pca, group = log_data$kelp_targeted_total,
         obs.scale = 1, var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_continuous(na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Kelp Targeted \nFish Biomass \nMean Total Annual")

ggbiplot(pca, group = log_data$kelp_targeted_total,
         obs.scale = 1, var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_continuous(na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Kelp Targeted \nFish Biomass \nMean Total Annual")

ggbiplot(pca_reduced, 
         group = log_data$kelp_targeted_total,
         obs.scale = 1, 
         var.scale = 1,) +
  theme_minimal() +
  scale_colour_continuous(na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "Kelp Targeted \nFish Biomass \nMean Total Annual")

ggbiplot(pca_reduced, 
         group = log_data$kelp_targeted_total,
         obs.scale = 1, 
         var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_colour_continuous(na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "Kelp Targeted \nFish Biomass \nMean Total Annual")

ggbiplot(pca_reduced, 
         group = log_data$kelp_targeted_total,
         obs.scale = 1, 
         var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_colour_continuous(na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "Kelp Targeted \nFish Biomass \nMean Total Annual")
```

### CCFRP

```{r ccfrp-targeted-total}
ggbiplot(pca, group = log_data$ccfrp_targeted_total,
         obs.scale = 1, var.scale = 1) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_continuous(na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "CCFRP Targeted \nFish Biomass \nMean Total Annual")


ggbiplot(pca, group = log_data$ccfrp_targeted_total,
         obs.scale = 1, var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_continuous(na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "CCFRP Targeted \nFish Biomass \nMean Total Annual")

ggbiplot(pca, group = log_data$ccfrp_targeted_total,
         obs.scale = 1, var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_continuous(na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "CCFRP Targeted \nFish Biomass \nMean Total Annual")

ggbiplot(pca_reduced, 
         group = log_data$ccfrp_targeted_total,
         obs.scale = 1, 
         var.scale = 1,) +
  theme_minimal() +
  scale_colour_continuous(na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "CCFRP Targeted \nFish Biomass \nMean Total Annual")

ggbiplot(pca_reduced, 
         group = log_data$ccfrp_targeted_total,
         obs.scale = 1, 
         var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_colour_continuous(na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "CCFRP Targeted \nFish Biomass \nMean Total Annual")

ggbiplot(pca_reduced, 
         group = log_data$ccfrp_targeted_total,
         obs.scale = 1, 
         var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_colour_continuous(na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "CCFRP Targeted \nFish Biomass \nMean Total Annual")
```
### Deep Reef Targeted

```{r deep-reef-targeted-total}
ggbiplot(pca, group = log_data$deep_reef_targeted_total,
         obs.scale = 1, var.scale = 1) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_continuous(na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Deep Reef Targeted \nFish Biomass \nMean Total Annual")

ggbiplot(pca, group = log_data$deep_reef_targeted_total,
         obs.scale = 1, var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_continuous(na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Deep Reef Targeted \nFish Biomass \nMean Total Annual")

ggbiplot(pca, group = log_data$deep_reef_targeted_total,
         obs.scale = 1, var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_x_continuous(limits = c(-10, 10))+
  scale_y_continuous(limits = c(-7,7))+
  scale_colour_continuous(na.value = NA) +
  labs(title = "All Trait Variables (log + 1 tranformed)",
       color = "Deep Reef Targeted \nFish Biomass \nMean Total Annual")

ggbiplot(pca_reduced, group = log_data$deep_reef_targeted_total, 
         obs.scale = 1, var.scale = 1,) +
  theme_minimal() +
  scale_colour_continuous(na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "Deep Reef Targeted \nFish Biomass \nMean Total Annual")

ggbiplot(pca_reduced, group = log_data$deep_reef_targeted_total,
         obs.scale = 1, var.scale = 1, var.axes = F) +
  theme_minimal() +
  scale_colour_continuous(na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "Deep Reef Targeted \nFish Biomass \nMean Total Annual")

ggbiplot(pca_reduced, group = log_data$deep_reef_targeted_total,
         obs.scale = 1, var.scale = 1, var.axes = F, labels = log_data$name) +
  theme_minimal() +
  scale_colour_continuous(na.value = NA) +
  labs(title = "Reduced Trait Variables",
       color = "Deep Reef Targeted \nFish Biomass \nMean Total Annual")
```



# NMDS ?

NMDS can have some value over PCA because it uses rank orders rather than absolute distance, so can also use different types of data.

```{r}
# nmds_data <-  data %>% 
#     select(name, mpa_class, four_region_north_ci,
#          age_yr,
#          distance_to_port:max_depth_m, 
#          asbs_overlap:hardened_armored_shore_km, 
#          citizensci_inaturalist_obs, 
#          total_hard_substrate:prop_soft) %>% 
#   select(!max_kelp_canopy_landsat_km2) %>% # Drop because too many NAs 
#   select(!larval_connectivity)  # Drop because too many NAs
#   
# nmds <- metaMDS(nmds_data[4:42], k = 2)
```


# References

Useful guide to PCA analyses: https://www.analyticsvidhya.com/blog/2016/03/pca-practical-guide-principal-component-analysis-python/
https://www.datacamp.com/tutorial/pca-analysis-r

Explanation for whether to reduce variables before doing PCA:
https://stats.stackexchange.com/questions/50537/should-one-remove-highly-correlated-variables-before-doing-pca


