---
title: "Habitat Diversity Analyses"
author: "Cori Lopazanski"
date: "3/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(janitor)
library(here)

share_path <- "/home/shares/ca-mpa/data/sync-data"
```

# About

This document contains the code for analyzing the questions related to habitat and functional diversity.

# Data

- 
- 

# MPA-Level Habitat Characteristics

```{r}
habitat_raw <- read_excel(here::here(share_path, "mpa-attributes.xlsx"),  na = c("NA", "."), trim_ws = T)
```

## Cleaning

```{r}
habitat <- habitat_raw %>% 
  clean_names() %>% 
  mutate(bioregion = recode_factor(bioregion,
                                   "NorCal"="Northern",
                                   "CenCal"="Central",
                                   "SoCal"="Southern")) %>%
  mutate(total_key_target_hab = select(., estuary_key_target_hab:submarine_canyons_all_depths_key_target_hab) %>% 
           rowSums(.))
  
```



## General Exploration

Proportion of "key target habitats" in the MPA

```{r}
ggplot(habitat) +
  geom_histogram(aes(x = total_key_target_hab, fill = designation)) +
  scale_x_continuous(n.breaks = 9) +
  scale_y_continuous(limits = c(0, 20), expand = c(0,0)) +
  facet_wrap(~bioregion) +
  labs(x = "total number of key target habitats within the MPA", 
       y = "frequency") +
  theme_classic()

ggplot(habitat) +
  geom_point(aes(x = size_km2, y = total_key_target_hab))
```

Distribution of different habitat types across the network

```{r}
habitat_long <- habitat %>% 
  select(name:max_kelp_canopy_cdfw_km2) %>% 
  pivot_longer(cols = sandy_beach_km:max_kelp_canopy_cdfw_km2,
               names_to = "habitat_type",
               values_to = "area") %>% 
  mutate(hab_prop = area/size_km2,
         hab_binary = if_else(hab_prop == 0, 0, 1))

# MPA table
site_table <- habitat_long %>% 
  pivot_wider(id_cols = name, 
              names_from = habitat_type,
              values_from = hab_binary) %>% 
  mutate(total_habitat = select(., sandy_beach_km:max_kelp_canopy_cdfw_km2) %>% rowSums(.)) %>% 
  full_join(. , habitat %>% select(name, monitoring_rov, monitoring_kelp, monitoring_intertidal)) %>% 
  rename(beach = sandy_beach_km,
         intertidal = rocky_inter_km,
         offshore_reef = offshore_rock_km,
         kelp = max_kelp_canopy_cdfw_km2)

# Join with monitoring yes/no
# ???
  
ggplot(data = habitat_long) +
  geom_point(aes(y = name, x = area, fill = habitat_type)) +
  facet_wrap(~bioregion)
```

# MPA-Level Monitoring Sites/Years

```{r}
monitor_raw <- read_excel(here::here(share_path, "ltm-site-table-cleaned.xlsx"))
```

## Create Site Table

```{r}
monitor <- monitor_raw %>% 
  clean_names() %>% 
  filter(mlpa_region == "SOUTH")

ggplot(data = monitor) +
  geom_point(aes(
```


To do:
- Get the ltm-site-table in the raw data folder; sync that here
- Check in with Julien about the sync 
- Do we have habitat data for the reference sites? Or best to assume those via monitoring efforts? That's problematic




