---
title: "means-map"
author: "Julien Brun"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'means_maps.html'))})
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(purrr)
library(sf)
library(tmap)
library(RColorBrewer)
```

## Reading the data

```{r paths}

# Define the paths and file names

path_aurora <- "/home/shares/ca-mpa/data/sync-data/"
shp_dir <- "spatial_data"
cencoos_dir <- "CeNCOOS_shapefiles/Ecological_Monitoring_Sites"
shp_path <- file.path(path_aurora, shp_dir)

shp_ca <- file.path(path_aurora, shp_dir, "ca_boundary_wgs84.shp")


input_file <- "Ecol_perform_metrics_means_working.xlsx" 

map_file <- "mpa-attributes.xlsx"


```


# Read the input files


```{r read inputs, message=FALSE, warning=FALSE}

perform_metrics <- readxl::read_excel(file.path(path_aurora, input_file), 
                            sheet=1, na = c("NA", "na"), trim_ws = T)

# map_attributes <-  readxl::read_excel(file.path(data_path, map_file), 
#                             sheet=1, na = c("NA", "."), trim_ws = T)

# Read California border
ca <- st_read(shp_ca)
st_crs(ca)


# st_perform_metrics <- perform_metrics %>% 
#   st_as_sf(.,coords=c("lon_wgs84", "lat_wgs84"), crs=4326)



```
We have some sites without coordinates. Also the precision of the location seems to vary betwee sites. **What should be our reference file?**

```{r, eval=FALSE}
perform_metrics %>% 
  filter(is.na(lat_wgs84) | is.na(lon_wgs84)) %>% 
  distinct(affiliated_mpa)
```
`gull island smr`, `carmel bay smca`, `anacapa island smr` are missing lat/long info. Dropping for now.


Still fails. There are character in the longitude column. Let's find which rows are affected

```{r}
perform_metrics %>% 
  filter(is.na(as.numeric(lon_wgs84))) %>% 
  distinct(affiliated_mpa)
```
`bodega head smca` seems to have some strange values for the longitude. Dropping for now.

```{r}
perform_metrics_cleaned <- perform_metrics %>% 
  filter(!is.na(lat_wgs84) | !is.na(lon_wgs84)) %>% 
  filter(!is.na(as.numeric(lon_wgs84))) 


# perform_metrics_static <- perform_metrics_cleaned %>%
#   select(group,mlpa_region, affiliated_mpa, mpa_class, mpa_designation, lat_wgs84, lon_wgs84) %>%
#   distinct()
# 
# dim(perform_metrics_static)

# st_perform_metrics <- perform_metrics %>% 
  # st_as_sf(.,coords=c("lon_wgs84", "lat_wgs84"), crs=4326)

```

```{r}

st_perform_metrics <- perform_metrics_cleaned %>% 
  st_as_sf(.,coords=c("lon_wgs84", "lat_wgs84"), crs=4326)


tm_reg <- tm_shape(ca) + 
  tm_borders("black", lwd = .5) +
tm_shape(st_perform_metrics) + 
  tm_symbols(col = "mlpa_region",  alpha = 0.8, border.lwd = 0, size = .2, title.col = "Main Regions")

tm_reg
```
Export as shapefile
```{r}
# Write combined shapefile
st_write(st_perform_metrics, dsn = file.path(path_aurora, cencoos_dir, "most_sampling_sites.geojson"), delete_dsn =TRUE)
```

Manually select the Island in QGIS and add a `channel_island` column coded with `1` (is a channel island site) and `0`.

```{r}

# read the new file in
st_perform_metrics_ci <- st_read(dsn = file.path(path_aurora, cencoos_dir, "most_sampling_sites_island.geojson"))

# Update the mlpa_region using this flag
st_perform_metrics_ci <- st_perform_metrics_ci %>%
  mutate(mlpa_region = ifelse(channel_island==1, "island", mlpa_region)) %>%
  select(-channel_island)


tm_reg4 <- tm_shape(ca) + 
  tm_borders("black", lwd = .5) +
tm_shape(st_perform_metrics_ci) + 
  tm_symbols(col = "mlpa_region",  alpha = 0.8, border.lwd = 0, size = .2, title.col = "Main Regions")

tm_reg4

# save the file
st_write(st_perform_metrics_ci,file.path(path_aurora, "Ecol_perform_metrics_means_4regions.csv"), delete_dsn =TRUE)


```


```{r clean data}
# Create missing fields

# estimate site status
habitat_sites_list$marine <- habitat_sites_list$marine %>%
  mutate(reserve_st = ifelse(mpa_design == "reference", "REF", "MPA"))

# Region acronym
habitat_sites_list$marine <- habitat_sites_list$marine %>%
  mutate(MPA_REGION = case_when( mpa_region == "North Central Coast" ~ "NCCSR",
                                 mpa_region == "Central Coast" ~ "CCSR",
                                 mpa_region == "South Coast" ~ "SCSR"
                                 )
    )

# Start and en year
habitat_sites_list$marine <- habitat_sites_list$marine %>%
  mutate(start_year = year(start_date),
         end_year = year(end_date)
         )




# Merge into one spatial dataframe
habitat_sites_all <- habitat_sites_list %>%
  bind_rows(.id = "source")


# Create habitat column from data source
habitat_sites_all <- habitat_sites_all %>%
  mutate(habitat = case_when( source == "pisco" ~ "kelp",
                              source == "marine" ~ "rocky",
                              source == "ccfrp" ~ "sandy"
                              )
         )

# Order the regions from North to South using factors
habitat_sites_all <- habitat_sites_all %>%
  mutate(mpa_reg_fact = factor(MPA_REGION, ordered = TRUE,
                                levels = c("NCSR", "NCCSR", "CCSR", "SCSR")
                               )
         )

# Compute the number of years each site has been active
habitat_sites_all <- habitat_sites_all %>%
  mutate(time_period = end_year - start_year + 1)

# Write combined shapefile
# st_write(habitat_sites_all, file.path(shp_path, "all_sampling_sites.shp"))

# Here I used QGIS to spatially select all the sites that were located around the channel island. I created a new column `chanIsland` where 1 means channel island site and 0 means coastal sites.

# I reread the data in
# habitat_sites_all <- st_read(file.path(shp_path, "all_sampling_sites.shp"))

```


```{r join}
# Join the shapefile and the means data frames

# Means_t come from Cori's code  `means-explore.Rmd`

habitat_sites_all_means <- left_join(habitat_sites_all, means_t, c("MPA_NAM"="affiliated_mpa"))

```



```{r map, message = FALSE}
# tmap_mode("plot")

tt <- habitat_sites_all_means %>% 
  filter(indicator %in% c("shannon", "Diversity")) %>% 
  filter(variable %in% c("INVALG", "All Fish")) %>%
  filter(year > 2015 & year < 2020) %>% 
  group_by(MPA_NAME, group) %>%
  summarise(mean_logRR = mean(log_RR))

tm_reg <- tm_shape(ca) + 
  tm_borders("black", lwd = .5) +
tm_shape(tt) + 
  tm_symbols(col = "mean_logRR",  alpha = 0.5, border.lwd = 0, palette = "-viridis") +
  tm_facets(by=c("group"))

tm_reg


# tmap_save(tm_reg, filename = "figs/mpa_regions.png")

```
```

