---
title: "Taxonomy Overlap"
author: "Cori Lopazanski"
date: '2022-05-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup 
```{r}
data_path <- "/home/shares/ca-mpa/data/sync-data/monitoring/taxonomy_tables"

library(tidyverse)
library(janitor)
library(here)
library(kableExtra)

```

# Read Taxonomy Tables

```{r}
ccfrp <- readxl::read_excel(here::here(data_path, "CCFRP_Taxonomy.xlsx"), sheet = 2,na = "NA") %>% 
  clean_names() %>% 
  mutate(group = "ccfrp")

dr <- readxl::read_excel(here::here(data_path, "DeepReef-ROV-Taxonomy.xlsx"), na = "NA") %>% 
  clean_names() %>% 
  mutate(group = "deep_reef")

rocky_bio <- readxl::read_excel(here::here(data_path, "RockyIntertidal-Biodiversity-Taxonomy.xlsx"), na = "NA") %>% 
  clean_names() %>% 
  mutate(group = "rocky_bio")

rocky_lt <- readxl::read_excel(here::here(data_path, "RockyIntertidal-LongTerm-Taxonomy.xlsx"), na = "NA") %>% 
  clean_names() %>% 
  mutate(group = "rocky_lt")

kf <- readxl::read_excel(here::here(data_path, "Kelp-Taxonomy.xlsx"), na = "NA") %>% 
  clean_names() %>% 
  mutate(group = "kelp_forest")

sb_seine <- read_csv(here::here(data_path, "SandyBeach-SurfZone-FishSeine-Taxonomy.csv"), na = "NA") %>% 
  clean_names() %>% 
  mutate(group = "sandy_beach_seine")

sb_bruv  <- read_csv(here::here(data_path, "SandyBeach-SurfZone-Bruv-Taxonomy.csv"), na = "NA") %>% 
  clean_names() %>% 
  mutate(group = "sandy_beach_bruv") 

```
# High-Level Cleaning

## General
```{r}
targeted_table_names <- c("group", "species_code", "scientific_name", "targeted")
```


## CCFRP
```{r}
ccfrp_clean <- ccfrp %>% 
  rename(targeted = fished) %>% 
  mutate(targeted = if_else(is.na(targeted) | targeted == "?", "NoData", targeted)) 

```

## Deep Reef
```{r}
dr_clean <- dr %>% 
  rename(species_code = pisco_code) %>% 
  mutate(targeted = if_else(is.na(targeted), "NoData", targeted)) 

```

## Kelp Forest
```{r}
kf_clean <- kf %>% 
   rename(species_code = pisco_classcode,
         scientific_name = scientific_name_accepted) %>% 
  mutate(targeted = if_else(is.na(targeted), "NoData", targeted),
         targeted = if_else(targeted == "Nontargeted", "Non-targeted", targeted))

```

## Rocky Intertidal
```{r}
rocky_bio_clean <- rocky_bio %>% 
  rename(scientific_name = species) %>% 
  mutate(targeted = NA)

rocky_lt_clean <- rocky_lt %>% 
  rename(species_code = marine_species_code,
         scientific_name = species)%>% 
  mutate(targeted = NA)

ri_clean <- full_join(rocky_bio_clean, rocky_lt_clean)
```

## Sandy Beach
```{r}
sb_bruv_clean <- sb_bruv %>% 
    rename(scientific_name = species_definition) %>% 
    mutate(targeted = if_else(is.na(targeted), "NoData", targeted))

sb_seine_clean <- sb_seine %>% 
  rename(scientific_name = scientific_name_accepted) %>% 
  filter(!is.na(scientific_name)) %>% 
  mutate(targeted = if_else(is.na(targeted), "NoData", targeted)) 

sb_clean <- full_join(sb_bruv_clean, sb_seine_clean)

sb_clean$targeted[sb_clean$targeted == "Nontargeted"] <- "Non-targeted"

```


# Goal 1: Create table which codes species as targeted and non-targeted

## Select columns from each individual taxonomy table and join together 
```{r}
tar <- rbind(ccfrp_clean[targeted_table_names],
                   kf_clean[targeted_table_names],
                   sb_clean[targeted_table_names],
                   dr_clean[targeted_table_names]) %>% 
  distinct(scientific_name, group, .keep_all = TRUE)  # remove duplicates
```

## Widen Table by Scientific Name
```{r}
tar_overlap <- tar %>%
  pivot_wider(id_cols = scientific_name, 
              names_from = group,
              values_from = targeted) %>% 
  replace(.=="NULL", NA) %>% 
  arrange(scientific_name)

```

## Export Table
```{r}
options(knitr.kable.NA = "")

names(tar_overlap) <- c("Scientific Name", "CCFRP", "Kelp Forest",
        "Sandy Beach (BRUV)", "Sandy Beach (Seine)", "Deep Reef") 

kbl(tar_overlap) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F, fixed_thead = T) %>% 
  save_kable(here::here("subteam-analyses",
                        "ltm-overlap",
                        "targeted-species-table.html"))
```


# Goal 2: Create one Species Trait Table 

## Identify Common Variables

```{r}

```





