---
title: "2-NMDS-Habitat-Log"
author: "Cori Lopazanski"
date: '2022-08-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

## Set Up

```{r}
# Packages
library(tidyverse)
library(vegan)

# Directories
int.dir <- file.path(getwd(), "analyses", "7habitat", "intermediate_data")

# Read MPA attribute subsets
att_data   <- readRDS(file.path(int.dir, "mpa_attributes_nmds_subset.Rds"))
att_noest  <- readRDS(file.path(int.dir, "mpa_attributes_nmds_subset_no_estuaries.Rds"))
att_meta   <- readRDS(file.path(int.dir, "mpa_attributes_nmds_metadata.Rds"))
att_labels <- readRDS(file.path(int.dir, "mpa_attributes_labels.Rds"))
```

## Build data

### Excluded MPAs (contain NAs)

```{r}
# Extract rows with any NA values
att_na <- att_data[rowSums(is.na(att_data)) > 0, ]

# List of excluded MPAs (contain NA)
rownames(att_na)

# Remove from attribute subsets
att_data <- att_data %>% 
  drop_na()

att_noest <- att_noest %>% 
  drop_na()
```

### Log Transformation

```{r}
att_data  <- log1p(att_data)
att_noest <- log1p(att_noest)
```


## All MPAs

### metaMDS (with z scores)

```{r}
mds <- metaMDS(att_data, distance = "euclidean", 
               trymax = 1500, maxit = 1500, 
               autotransform = F, trace = F)
mds
stressplot(mds)
```

### envfit for vectors

Determine the relative contribution of variables to the separation

```{r}
# Output of metaMDS and the original matrix
fit <- envfit(mds, att_data, perm = 999)
```

### build plot dfs

```{r results='hide', fig.show='hide'}
# Extract pvalues for each attribute
fit_pval <- fit$vectors$pvals %>% 
  as.data.frame() %>% 
  rownames_to_column("attribute") %>% 
  dplyr::rename("pvals" = ".")

# Get coordinates for each attribute vector
fit_att <- fit %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("attribute") %>% 
  full_join(., fit_pval, by = "attribute") %>% 
  mutate(sig = if_else(pvals == 0.001, "yes", "no")) %>% 
  left_join(., att_labels)

# Get coordinates for each MPA (NMDS dimension)
nmds_df <- scores(mds, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("name") %>% 
  left_join(att_meta, by = "name") %>% 
  mutate(region = as.factor(region))

# Get coordinates for ellipses
plot.new()
ell <- ordiellipse(mds, nmds_df$region, kind = "se", conf = 0.95, label = T)

ell_df <- data.frame()
for(i in levels(nmds_df$region)){
  ell_df <- rbind(ell_df, 
                  cbind(as.data.frame(with(nmds_df[nmds_df$region==i,],
                                           vegan:::veganCovEllipse(ell[[i]]$cov, 
                                                                   ell[[i]]$center, 
                                                                   ell[[i]]$scale))), 
                          region = i))}


# Determine scale factor for vectors
scale_factor <- 4
```


### plot 

#### all

```{r}
ggplot(nmds_df, aes(x = NMDS1, y = NMDS2)) +
  # Points
  geom_point(aes(color = region, shape = region),
             size = 3, alpha = 0.5) +
  # Ellipses with ordiellipse
  geom_path(data = ell_df, 
            aes(x = NMDS1, y = NMDS2, color = region), 
            linetype = 5) + 
  # Vectors
  geom_segment(data = fit_att, 
               aes(x = 0, xend = NMDS1*scale_factor,
                   y = 0, yend = NMDS2*scale_factor),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  #  Vector Labels
  geom_text(data = fit_att, 
            aes(x = NMDS1*(scale_factor + 0.5),
                y = NMDS2*(scale_factor + 0.5),
                label = att_short), size = 2.5) +
  # Formatting
  coord_fixed() +
  labs(title = "NMDS",
       color = "",
       shape = "") + 
  theme_classic()
```

#### centroids and ellipses

```{r}
# Get cooordinates for centroids
centroids <- nmds_df %>% 
  group_by(region) %>% 
  summarize(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))

ggplot(nmds_df, aes(x = NMDS1, y = NMDS2)) +
  # Centroids
  geom_point(data = centroids, aes(color = region)) +
  # Ellipses with ordiellipse
  geom_path(data = ell_df, 
            aes(x = NMDS1, y = NMDS2, color = region), 
            linetype = 5) + 
  # Formatting
  coord_fixed() +
  labs(title = "NMDS",
       color = "",
       shape = "") + 
  theme_classic()
```


## Without Estuaries

### metaMDS

```{r}
mds_noest <- metaMDS(att_noest, distance = "euclidean", 
                     maxit = 999, trymax = 999, autotransform = F, trace = F)
mds_noest
stressplot(mds_noest)
```

### envfit for vectors

determine relative contribution of variables to the separation

```{r}
# Output of metaMDS and the original matrix
fit_noest <- envfit(mds_noest, att_noest, perm = 999)
```


### build plot dfs

```{r results='hide', fig.show='hide'}
# Extract pvalues for each attribute
fit_pval_noest <- fit_noest$vectors$pvals %>% 
  as.data.frame() %>% 
  rownames_to_column("attribute") %>% 
  dplyr::rename("pvals" = ".")

# Get coordinates for each attribute vector
fit_att_noest <- fit_noest %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("attribute") %>% 
  full_join(., fit_pval_noest, by = "attribute") %>% 
  mutate(sig = if_else(pvals == 0.001, "yes", "no")) %>% 
  left_join(., att_labels)

# Get coordinates for each MPA (NMDS dimension)
nmds_df_noest <- scores(mds_noest, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("name") %>% 
  left_join(att_meta, by = "name") %>% 
  mutate(region = as.factor(region))

# Get coordinates for ellipses
plot.new()
ell_noest <- ordiellipse(mds_noest, nmds_df_noest$region, kind = "se", conf = 0.95, label = T)

ell_df_noest <- data.frame()
for(i in levels(nmds_df_noest$region)){
  ell_df_noest <- rbind(ell_df_noest, 
                  cbind(as.data.frame(with(nmds_df_noest[nmds_df_noest$region==i,],
                                           vegan:::veganCovEllipse(ell_noest[[i]]$cov, 
                                                                   ell_noest[[i]]$center, 
                                                                   ell_noest[[i]]$scale))), 
                          region = i))}

# Determine scale factor for vectors
scale_factor_noest <- 4
```

### plot


#### all

```{r}
ggplot(nmds_df_noest, aes(x = NMDS1, y = NMDS2)) +
  # Points
  geom_point(aes(color = region, shape = region),
             size = 3, alpha = 0.5) +
  # Ellipses with ordiellipse
  geom_path(data = ell_df_noest, 
            aes(x = NMDS1, y = NMDS2, color = region), 
            linetype = 5) + 
  # Vectors
  geom_segment(data = fit_att_noest, 
               aes(x = 0, xend = NMDS1*scale_factor_noest,
                   y = 0, yend = NMDS2*scale_factor_noest),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  # Vector Labels
  geom_text(data = fit_att_noest, 
            aes(x = NMDS1*(scale_factor_noest+0.5),
                y = NMDS2*(scale_factor_noest+0.5),
                label = att_short), size = 2.5) +
  # Formatting
  coord_fixed() +
  labs(title = "NMDS",
       color = "",
       shape = "") + 
  theme_classic()
```

#### centroids and ellipses

```{r}
# Get cooordinates for centroids
centroids_noest <- nmds_df_noest %>% 
  group_by(region) %>% 
  summarize(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))

ggplot(nmds_df_noest, aes(x = NMDS1, y = NMDS2)) +
  # Centroids
  geom_point(data = centroids_noest, aes(color = region)) +
  # Ellipses with ordiellipse
  geom_path(data = ell_df_noest, 
            aes(x = NMDS1, y = NMDS2, color = region), 
            linetype = 5) + 
  # Formatting
  coord_fixed() +
  labs(title = "NMDS",
       color = "",
       shape = "") + 
  theme_classic()
```

### test for regional differences 

#### permutest

test for differences in dispersion 

```{r}
# Distance Matrix
dismat <- vegdist(att_noest, method = "euclidean")

# disper
disper <- betadisper(dismat, type = "centroid", group = att_meta$region[att_meta$name %in% rownames(att_noest)])
plot(disper)

# Test for differences in dispersion for the regions
test <- permutest(disper)
test
```

#### permanova

test for differences among regions

```{r}
# Difference among groups (regions)
permanova <- adonis2(formula = dismat ~ region, 
                     mwrhos = "euclidean",
                    data = att_meta[att_meta$name %in% rownames(att_noest),])

permanova

hsd <- TukeyHSD(disper)
hsd
```

