---
title: "Kelp Forest Communities"
author: "Cori Lopazanski"
date: '2022-07-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

## Setup

```{r}
# Packages
library(tidyverse)
library(vegan)
library(ggbiplot)
library(naniar)

# Directories
int.dir <- file.path("analyses", "7habitat","intermediate_data")

# Read Data
## Attributes
att_data <- readRDS(file.path(int.dir, "mpa_attributes_nmds_subset.Rds"))

## Attribute metadata
att_meta <- readRDS(file.path(int.dir, "mpa_attributes_nmds_metadata.Rds"))

## Kelp Fish Counts
kf_fish <- readRDS(file.path(int.dir, "species_matrix_kelp_fish_1620.Rds")) 

## Kelp Swath Counts
#kf_swath <- readRDS(file.path(int.dir, "species_matrix_kelp_swath_1620.Rds"))

# MPA Coordinates
mpas <- readRDS(file.path(int.dir, "mpa_coords.Rds")) %>% 
  select(name, lat_dd) %>% 
  mutate(name = str_replace(name, " \\s*\\([^\\)]+\\)", ""))

rownames(mpas) <- NULL
```

## Fish Community

Species matrix for 2016-2020 average count per year.

```{r}
# Get MPA attributes for MPAs in kelp species matrix 
kf_env <- att_data %>% 
  filter(rownames(.) %in% kf_fish$name) 

# Save metadata df
kf_md <- att_meta %>% 
  select(name, region) %>% 
  filter(name %in% rownames(kf_env))

# Set up species matrix for vegan analyses
kf_fish[is.na(kf_fish)] <- 0
kf_fish <- kf_fish %>% column_to_rownames("name")
kf_env[is.na(kf_env)] <- 0

# Drop two species: BAIT (baitball) and ATHE (atherinidae)
# Following recommendations by P.R. and M.C.
kf_fish <- kf_fish %>% 
  select(-BAITBALL, -ATHE)

# Get latitude
mpa_lat <- mpas %>% 
  mutate(name = str_to_lower(name)) %>% 
  column_to_rownames("name")

# Join latitude to env data
kf_env <- merge(kf_env, mpa_lat, by = "row.names") %>% 
  column_to_rownames("Row.names")

```

### examine richness and diversity

```{r}
# kf fish richness
kf_sppr <- specnumber(kf_fish)

# difference in richness between regions?
kf_sppr_aov <- aov(kf_sppr ~ region, data = kf_md)
summary(kf_sppr_aov) # no diff

# kf fish diversity
kf_shan <- diversity(kf_fish)

# difference in diversity between regions?
kf_shan_aov <- aov(kf_shan ~ region, data = kf_md)
summary(kf_shan_aov) # no diff

# difference in species composition between regions?
kf_perm <- adonis2(kf_fish ~ region, data = kf_md)
kf_perm # no diff
```

### nmds 

```{r}
# nmds of fish community (raw)
kf_mds <- metaMDS(kf_fish, autotransform = F)
kf_mds
stressplot(kf_mds)

# plot nmds of fish community
kf_df <- scores(kf_mds, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("name") %>% 
  full_join(kf_md, by = "name")

kf_plot <- ggplot(kf_df,
                  aes(x = NMDS1, y = NMDS2,
                      color = region)) +
  geom_point(size = 3, alpha = 0.5) +
  labs(title = "KF Fish Community NMDS - RAW") + 
  theme_classic()
kf_plot

# nmds of fish community (log)
kf_fish_log <- kf_fish %>% mutate_if(., is.numeric, log1p)
kf_mds_log <- metaMDS(kf_fish_log, autotransform = F)
kf_mds
stressplot(kf_mds)

```



### envfit: species driving separation?

```{r}
# envfit - how do species contribute to dissimilarity
kf_fit <- envfit(kf_mds, kf_fish, perm = 999)

kf_fit_pvals <- kf_fit$vectors$pvals %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  dplyr::rename("pvals" = ".") # over half of them have = 0.001

```

### bioenv: link environmental variables to community dissimilarity

```{r}
# bioenv how do env vars contribute to dissimilarity
kf_bioenv <- bioenv(comm = kf_fish, env = kf_env, metric = "euclidean")
kf_bioenv
summary(kf_bioenv)

# select best model
kf_model <- kf_bioenv$models[[kf_bioenv$whichbest]] 
kf_keep <- kf_model$best

# use envfit to get vectors for top model variables
kf_envfit <- envfit(kf_mds, kf_env[, kf_keep], perm = 999, na.rm = TRUE)

# extract coordinates for each attribute
kf_envfit_coords <- kf_envfit %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("attribute") %>% 
  mutate(att.label = c("soft 200-3000", "surfgrass", "eelgrass", "coastal marsh"))

# plot
ggplot(kf_df, aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(color = region,
                 shape = region),
             size = 3, alpha = 0.8) +
  geom_segment(data = kf_envfit_coords, 
               aes(x = 0, xend = NMDS1*10,
                   y = 0, yend = NMDS2*10),
               arrow = arrow(length = unit(0.15, "cm")),
               col = "black") +
  geom_text(data = kf_envfit_coords, 
            aes(label = att.label), 
            size = 2.5) +
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +
  labs(title = "KF Fish",
       color = "",
       shape = "") + 
  theme_classic()
```

### pca of kf_env

```{r}
# pca
pca <- prcomp(kf_env, scale = T)
summary(pca)

# retain
pca_scores <- scores(pca, display = "sites")

pca_retain <- pca_scores[,1:4] %>% as.data.frame()

# model
cca <- cca(kf_fish ~ ., data = pca_retain)

# test
anova(cca, by = "axis")

# plot
plot(cca)
```



## Swath (Algae/Inverts)

```{r}
# # Get MPA attributes for MPAs in kelp species matrix 
# ks_env <- att_data %>% 
#   filter(rownames(.) %in% kf_swath$name) 
# 
# # Save metadata df
# ks_md <- att_meta %>% 
#   select(name, region) %>% 
#   filter(name %in% rownames(ks_env))
# 
# # Set up species matrix for vegan analyses
# kf_swath[is.na(kf_swath)] <- 0
# kf_swath <- kf_swath %>% column_to_rownames("name")
# ks_env[is.na(ks_env)] <- 0
```

### examine richness and diversity

```{r}
# # kf swath richness
# ks_sppr <- specnumber(kf_swath)
# 
# # difference in richness between regions?
# ks_sppr_aov <- aov(ks_sppr ~ region, data = ks_md)
# summary(ks_sppr_aov) # no diff
# 
# # kf fish diversity
# ks_shan <- diversity(kf_swath)
# 
# # difference in diversity between regions?
# ks_shan_aov <- aov(ks_shan ~ region, data = ks_md)
# summary(ks_shan_aov) # no diff
# 
# # difference in species composition between regions?
# ks_perm <- adonis2(kf_swath ~ region, data = ks_md)
# ks_perm # no diff
```

### nmds

```{r}
# # nmds of invert/algae community
# ks_mds <- metaMDS(kf_swath, distance = "euclidean", trymax = 999)
# ks_mds
# stressplot(ks_mds)
# 
# # plot nmds of fish community
# ks_df <- scores(ks_mds, display = "sites") %>% 
#   as.data.frame() %>% 
#   rownames_to_column("name") %>% 
#   full_join(ks_md, by = "name")
# 
# ks_plot <- ggplot(ks_df,
#                   aes(x = NMDS1, y = NMDS2,
#                       color = region)) +
#   geom_point(size = 3, alpha = 0.5) +
#   labs(title = "KF Swath NMDS") + 
#   theme_classic()
# ks_plot
```

### envfit: species driving separation?

```{r}
# # envfit - how do species contribute to dissimilarity
# ks_fit <- envfit(ks_mds, kf_swath, perm = 999)
# 
# ks_fit_pvals <- ks_fit$vectors$pvals %>% 
#   as.data.frame() %>% 
#   rownames_to_column("species") %>% 
#   dplyr::rename("pvals" = ".") # over half of them have = 0.001

```

### bioenv: link environmental variables to community dissimilarity

```{r}
# # bioenv how do env vars contribute to dissimilarity
# ks_bioenv <- bioenv(comm = kf_swath, env = ks_env)
# ks_bioenv
# 
# # select best model
# ks_model <- ks_bioenv$models[[ks_bioenv$whichbest]] 
# ks_keep <- ks_model$best
# 
# # use envfit to get vectors for top model variables
# ks_envfit <- envfit(ks_mds, ks_env[, ks_keep], perm = 999, na.rm = TRUE)
# 
# # extract coordinates for each attribute
# ks_envfit_coords <- ks_envfit %>% 
#   scores(., display = "vectors") %>% 
#   as.data.frame() %>% 
#   rownames_to_column("attribute")
# 
# #plot
# ggplot(ks_df, aes(x = NMDS1, y = NMDS2)) +
#   coord_fixed() +
#   geom_point(aes(color = region,
#                  shape = region),
#              size = 3, alpha = 0.8) +
#   geom_segment(data = ks_envfit_coords,
#                aes(x = 0, xend = NMDS1,
#                    y = 0, yend = NMDS2),
#                arrow = arrow(length = unit(0.15, "cm")),
#                col = "black") +
#   geom_text(data = ks_envfit_coords, aes(label = attribute), size = 3) +
#   geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
#   geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +
#   scale_x_continuous(limits = c(-0.2, 0.25)) +
#   scale_y_continuous(limits = c(-0.05, 0.25)) +
#   labs(title = "KF Swath",
#        color = "",
#        shape = "") +
#   theme_classic()
```

