---
title: "Exploring Methods: Spatial Distribution of Catch"
author: "Cori Lopazanski"
date: "2023-08-01"
output: 
  html_document: 
    fig_width: 6
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r}
# Clear workspace
rm(list = ls())

# Packages
library(tidyverse)
library(janitor)
library(sf)
library(rfishbase)

# Directories
basedir <- "/Users/lopazanski/Library/CloudStorage/GoogleDrive-lopazanski@ucsb.edu/Shared drives/NCEAS MPA network assessment/MPA Network Assessment: Working Group Shared Folder/data/sync-data"
gisdir <- file.path(basedir, "gis_data/processed")
datadir <- "analyses/2performance_fisheries/analyses/blocks/data"
fishdir <- "/Users/lopazanski/Documents/github/nceas/CDFW-fishing-data"

```

# Read Data

```{r}
mpas <- readRDS(file.path(gisdir, "CA_MPA_polygons.Rds"))
mpa_meta <- readRDS(file.path(basedir, "mpa_traits", "processed", "mpa_attributes_general.Rds"))
mpa_defacto <- readRDS("data/mpa_traits/mpa_defacto.Rds") %>% 
  pivot_wider(names_from = group, values_from = mpa_class) %>% 
  mutate(defacto = if_else(if_any(c(kelp, rocky, beach, ccfrp, deep_reef, surf), ~.x == "SMCA"), "SMCA", "SMR"))

blocks <- readRDS(file.path(datadir, "blocks_by_mlpa_region_w_mpa_stats.Rds"))
landings_raw <- readRDS(file.path(fishdir, "CDFW_2000_2020_landings_receipts.Rds"))
fisheries <- readRDS(file.path(datadir, "species_fishery_group_key.Rds")) %>% 
  rename(species_id = spp_code_num)

contours <- read_sf("/Users/lopazanski/Documents/github/nceas/CDFW-fishing-data/contours_5m")

```

# Clean Data

## Clean up landings data

```{r}
landings <- landings_raw %>% 
  # Drop invalid blocks
  filter(!(block_type %in% c("Invalid"))) %>% 
  # Drop missing weights (6 obs) 
  filter(!is.na(landings_lb)) %>% 
  left_join(fisheries)

# Inspect missing values
colSums(is.na(landings))

# Key:
# Missing permits: ~ 2.4 million obs
# Missing price/value: 773 obs.
# Not included in current fishery groupings: 239,304 obs.

```

## Create simple blocks dataframe with MPA info

```{r}
patterns <- c("SMRMA", "SMR", "SMCA \\(No-Take\\)", "FMR", "SMP", "SMCA")

blocks_simple <- blocks %>% 
  select(block_id, mlpa_region:mpa_km2, geometry) %>% 
  mutate(mpa_types = str_extract_all(mpas, paste(patterns, collapse = "|"))) %>% 
  unnest(mpa_types) %>% 
  pivot_wider(names_from = "mpa_types", values_from = "mpa_types", values_fn = first) %>% clean_names() %>% 
  select(block_id:mpa_km2, smr, smca_no_take, FMR, smca, smrma, smp, geometry) %>% 
  mutate(highly = if_else(smr > 0 | smca_no_take > 0 | FMR > 0, "Contains defacto SMR", NA),
         highly = if_else(is.na(highly) & mpa_n > 0, "SMCA or lower", highly)) %>% 
  unite("mpa_types", smr:smp, sep = ", ", na.rm = T) %>% 
  mutate(mpa_types = if_else(mpa_types == "", NA, mpa_types))


```

### Plot
```{r}
# CA/OR, Alder Creek, Pigeon Point, Point Conception, CA/MEX 
region_lats <- c(42, 39.0, 37.18, 34.5, 32.5) %>% rev()

# Get land
usa <- rnaturalearth::ne_states(country="United States of America", returnclass = "sf")
foreign <- rnaturalearth::ne_countries(country=c("Canada", "Mexico"), returnclass = "sf")

# Theme
my_theme <-  theme(axis.text=element_text(size=7),
                   axis.text.y = element_text(angle = 90, hjust = 0.5),
                   axis.title=element_blank(),
                   legend.text=element_text(size=8),
                   legend.title=element_text(size=9),
                   # Gridlines
                   panel.grid.major = element_blank(), 
                   panel.grid.minor = element_blank(),
                   panel.background = element_blank(), 
                   axis.line = element_line(colour = "black"),
                   # Legend
                   legend.key.size = unit(0.4, "cm"),
                   legend.key = element_rect(fill=alpha('blue', 0)),
                   legend.background = element_rect(fill=alpha('blue', 0)))


# Plot data - MPA status based on regulation
g <- ggplot() +
  # Plot blocks
  geom_sf(data=blocks_simple %>% 
            filter(mlpa_region == "South Coast"), mapping=aes(geometry=geometry, fill=highly, alpha=mpa_yn), lwd=0.2) +
  # Plot region lines
  geom_hline(mapping=aes(yintercept=region_lats)) +
  # Plot land
  geom_sf(data=foreign, fill="grey80", color="white", lwd=0.3) +
  geom_sf(data=usa, fill="grey80", color="white", lwd=0.3) +
  # MPAs
  geom_sf(data = mpas, fill = "grey", alpha = 0.5, lwd = 0.2) +
  # Legend
  scale_fill_discrete(name="Regulation", na.translate = F) +
  scale_alpha_manual(name="Block type", values=c(0.9, 0.2)) +
  # Crop
  coord_sf(xlim = c(-121, -117), ylim = c(32.5, 34.45)) +
  # Theme
  theme_bw() + my_theme
g


```


```{r}
# Plot data - MPA status based on regulation
g <- ggplot() +
  # Plot blocks
  geom_sf(data=blocks_simple %>% 
            filter(mlpa_region == "South Coast") %>% 
            mutate(highly = if_else(mpa_km2 < 0.5, "MPA < 0.5 km2", highly)), 
          mapping=aes(geometry=geometry, fill=highly, alpha=mpa_yn), lwd=0.2) +
  # Plot region lines
  geom_hline(mapping=aes(yintercept=region_lats)) +
  # Plot land
  geom_sf(data=foreign, fill="grey80", color="white", lwd=0.3) +
  geom_sf(data=usa, fill="grey80", color="white", lwd=0.3) +
  # MPAs
  geom_sf(data = mpas, fill = "grey", alpha = 0.5, lwd = 0.2) +
  # Legend
  scale_fill_discrete(name="Regulation", na.translate = F) +
  scale_alpha_manual(name="Block type", values=c(0.9, 0.2)) +
  # Crop
  coord_sf(xlim = c(-121, -117), ylim = c(32.5, 34.45)) +
  # Theme
  theme_bw() + my_theme
g

```


# Build Data

## Calculate CPUE

We define CPUE as in Richerson et al. 2020 (<https://www.sciencedirect.com/science/article/pii/S016578362030045X>). CPUE is weight per [species or species group] per trip, where a trip is the total landings of the [species or species group] landed by a single vessel from a specific block on the same date in the same port.

1. Calculate total landings of each species (s) landed by an individual vessel (v) in a single block (b) using a specific gear (g) on a particular day (t), and landed in a particular port (p),
2. Create dummy variables for whether or not there is an MPA in the region on a particular day (t),
3. Create dummy variable for whether or not there is an MPA in the block on a particular day (t)

```{r}
data <- landings %>% 
  # Vessel - Block - Gear - Day - Species
  group_by(vessel_id, species_id, species, block_id, 
           gear_type, gear_id, gear, 
           port_complex, port_id, port,
           year, month, date) %>% 
  # Calculate total landings for each "trip"
  summarize(total_lbs = sum(landings_lb, na.rm = T)) %>% 
  ungroup() %>% 
  # Join the MLPA region and MPA status
  left_join(., blocks_simple) %>% 
  # Create dummy for whether or not there is an MPA in the region for each block
  mutate(mpa_in_region = case_when(date < '2007-09-01' & mlpa_region == "Central Coast" ~ 0,
                                 date < '2010-05-01' & mlpa_region == "North Central Coast" ~ 0,
                                 date < '2012-12-01' & mlpa_region == "North Coast" ~ 0,
                                 date < '2012-01-01' & mlpa_region == "South Coast" ~ 0,
                                 mlpa_region == "San Francicso Bay" ~ 0,
                                 TRUE ~ 1)) %>% 
  # Create dummy for whether or not there is an MPA (implemented) in the block
  mutate(mpa_in_block = case_when(mpa_yn == "MPA" & mpa_in_region == 1 ~ 1,
                                  TRUE ~ 0)) 
```


```{r}
## Explore Rockfish - Not Currently Seeming Correct
# data_rockfish <- data %>% 
#   filter(species_id %in% fisheries$species_id[fisheries$fishery == "rockfish"]) %>% 
#   group_by(vessel_id, block_id, gear, port_complex, port, year, month, date) %>% 
#   # Summarize across all rockfish species for a given trip-block
#   summarize(total_lbs = sum(total_lbs))
# 
# test_multiple_blocks <- data_rockfish %>% 
#   group_by(vessel_id, gear, port_complex, port, year, month, date) %>% 
#   summarize(total_lbs = sum(total_lbs, na.rm = T),
#             n_blocks = n()) 
# 
# length(test_multiple_blocks$n_blocks[test_multiple_blocks$n_blocks>1])
# 1683/134529
# 
# test_multiple_gears <- data_rockfish %>% 
#   group_by(vessel_id, year, month, date) %>% 
#   summarize(n_gears = length(unique(gear)))
# 
# length(unique(data$gear_id)) # 76 different gear configurations
# 
# length(test_multiple_gears$n_gears[test_multiple_gears$n_gears>1]) #1436 gears > 1 on single date
# 1436/131253
# 
# # Examine most common gear types
# gear_types <- data %>% 
#   group_by(gear_type, gear_id, gear) %>% 
#   summarize(receipts = n(), 
#             landings = sum(total_lbs, na.rm = T),
#             vessels = length(unique(vessel_id)))
# 
# # Plot common gears by year
# gear_landings <- data %>% 
#   group_by(gear_type, year) %>% 
#   summarize(landings = sum(total_lbs, na.rm = T)) 
# 
# ggplot(data = gear_landings) + 
#   geom_line(aes(x = year, y = landings, color = fct_reorder(gear_type, desc(landings)))) +
#   labs(color = NULL, 
#        x = "Year", y = "Total rockfish catch (lbs) per major gear type") +
#   theme_minimal()
```

## Calculate Standardized CPUE

Scale the CPUE to center around the mean and convert to unit variance based on the pre-implementation moments.

Since the network was implemented regionally, which pre-implementation mean do we want to use? One that is consistent before any MPA was implemented? Region-specific?

```{r}
# Start with calculating pre-implementation mean based on zero MLPA MPAs implemented
# e.g. prior to first region implemented (before 09-01-2007; Central)
# which is technically after Channel Islands MPAs and some other MPAs...
data_std <- data  %>% 
  filter(!(vessel_id %in% c(-1, 0))) %>% 
  # Calculate mean landings for each species for each vessel for each gear
  # in each block in a given year (across all trips)
  group_by(vessel_id, species, species_id, gear_type,
           block_id, mlpa_region, mpa_yn, mpa_in_region) %>% 
  summarize(mean_lbs = mean(total_lbs)) %>% 
  arrange(-vessel_id, block_id) %>% 
  pivot_wider(names_from = mpa_in_region, values_from = mean_lbs) %>% 
  rename(mean_lbs_pre = `0`, mean_lbs_post = `1`) %>% 
  filter(!(is.na(mean_lbs_pre) | is.na(mean_lbs_post))) %>%  # 225925 > 19476
  mutate(logRR = log(mean_lbs_post/mean_lbs_pre)) %>% 
  mutate(block_ord = fct_reorder(factor(block_id), logRR)) %>% 
  left_join(fisheries)

ggplot(data = data_std) +
  geom_point(aes(x = block_id, y = logRR, color = fishery))


mutate(pre_post = case_when(date < '2007-09-01' ~ "pre",
                              date >= '2007-09-01' ~ "post")) %>% 


# Calculate mean and standard deviation
data_pre.mean <- mean(data_pre$total_lbs)
data_pre.sd <- sd(data_pre$total_lbs)

# Create new cpue_std = (cpue - mean before impl.)/sd before impl.
data$cpue_std <- (data$total_lbs - data_pre.mean)/data_pre.sd

```

## Difference Pre/Post

```{r}
# Calculate the total standardized CPUE for each block pre-implementation
annual_cpue <- data %>% 
  group_by(block_id, mpa_yn, year) %>% 
  summarize(total_lbs = sum(total_lbs),
            annual_cpue = mean(cpue_std)) %>% 
  ungroup() %>% 
  group_by(block_id)

ggplot(data = annual_cpue) +
  geom_path(aes(x = year, y = total_lbs, group = block_id, color = mpa_yn))+
  scale_y_continuous(limits = c(0, 200000))

plot_pre <- data %>% 
  filter(date < '2007-09-01') %>% 
  group_by(block_id, year) %>% 
  summarize(annual_cpue = sum(cpue_std)) %>% 
  group_by(block_id) %>% 
  summarize(pre_cpue = mean(annual_cpue))

plot_post <- data %>% 
  filter(date > '2007-09-01') %>% 
  group_by(block_id, year) %>% 
  summarize(annual_cpue = sum(cpue_std)) %>% 
  group_by(block_id) %>% 
  summarize(post_cpue = mean(annual_cpue))

plot <- full_join(plot_pre, plot_post) 

plot$diff <- plot$post_cpue-plot$pre_cpue
```

## Version with Inshore Only

```{r}
# Inshore blocks only
inshore <- blocks %>% 
  filter(block_type != "Offshore")

# Subset data to only the inshore blocks
data_inshore <- data %>% 
  filter(block_id %in% inshore$block_id)


data_inshore_pre <- data_inshore %>% 
  filter(date < '2007-09-01')
  
data_inshore_pre.mean <- mean(data_inshore_pre$total_lbs)
data_inshore_pre.sd <- sd(data_inshore_pre$total_lbs)
data_inshore$cpue_std <- (data_inshore$total_lbs - data_inshore_pre.mean)/data_inshore_pre.sd
```

## Difference Pre/Post

```{r}
inshore_pre <-data_inshore %>% 
  filter(date < '2007-09-01') %>% 
  group_by(block_id) %>% 
  summarize(pre_cpue = sum(cpue_std))

inshore_post <- data_inshore %>% 
  filter(date > '2007-09-01') %>% 
  group_by(block_id) %>% 
  summarize(post_cpue = sum(cpue_std))

plot_inshore <- full_join(inshore_pre, inshore_post) 

plot_inshore$diff <- plot_inshore$post_cpue-plot_inshore$pre_cpue

plot_inshore2 <- plot_inshore %>% 
  mutate(test = scales::rescale(diff, to = c(-1, 1))) %>% 

```

m# Plots

## Theme Setup

```{r}
# CA/OR, Alder Creek, Pigeon Point, Point Conception, CA/MEX 
region_lats <- c(42, 39.0, 37.18, 34.5, 32.5) %>% rev()

# Get land
usa <- rnaturalearth::ne_states(country="United States of America", returnclass = "sf")
foreign <- rnaturalearth::ne_countries(country=c("Canada", "Mexico"), returnclass = "sf")

# Theme
my_theme <-  theme(axis.text=element_text(size=7),
                   axis.text.y = element_text(angle = 90, hjust = 0.5),
                   axis.title=element_blank(),
                   legend.text=element_text(size=8),
                   legend.title=element_text(size=9),
                   # Gridlines
                   panel.grid.major = element_blank(), 
                   panel.grid.minor = element_blank(),
                   panel.background = element_blank(), 
                   axis.line = element_line(colour = "black"),
                   # Legend
                   legend.key.size = unit(0.4, "cm"),
                   legend.key = element_rect(fill=alpha('blue', 0)),
                   legend.background = element_rect(fill=alpha('blue', 0)))
```

## All Blocks

```{r}
ggplot() +
  # Blocks
  geom_sf(data = blocks, lwd = 0.2) +
  # Landings 
  geom_sf(data = plot %>% 
            left_join(., blocks), aes(fill = diff, geometry = geometry)) +
  # MPAs
  geom_sf(data = mpas, color = "black", lwd = 0.2) +
  # Region lines
  geom_hline(mapping=aes(yintercept=region_lats)) +
  # Land
  geom_sf(data=foreign, fill="grey80", color="white", lwd=0.3) +
  geom_sf(data=usa, fill="grey80", color="white", lwd=0.3) +
  # Crop
  coord_sf(xlim = c(-124, -120), ylim = c(34.5, 37.2)) +
  # Color
  scale_fill_gradient2(midpoint = 0) +
  theme_bw() + my_theme + 
  labs(title = "Rockfish (Sebastes spp.)",
       fill = "Difference in std. CPUE")

#ggsave("rockfish_central.png", width = 4, height = 6, units = c("in"), dpi = 150)
```

## Inshore Blocks

```{r}
ggplot() +
  # Blocks
  geom_sf(data = blocks, lwd = 0.2) +
  # Landings 
  geom_sf(data = plot_inshore2 %>% 
            left_join(., blocks), aes(fill = test, geometry = geometry)) +
   # MPAs
  geom_sf(data = mpas, color = "black", lwd = 0.2) +
  # Region lines
  geom_hline(mapping=aes(yintercept=region_lats)) +
  # Land
  geom_sf(data=foreign, fill="grey80", color="white", lwd=0.3) +
  geom_sf(data=usa, fill="grey80", color="white", lwd=0.3) +
  # Crop
  coord_sf(xlim = c(-123, -120), ylim = c(34.5, 37.2)) +
  # Color
  scale_fill_gradient2(midpoint = 0) +
  theme_bw() + my_theme + 
  labs(title = "Rockfish (Sebastes spp.)",
       fill = "Standardized difference post-MPA")

```


```{r}
contours_500 <- contours %>% 
  filter(CONTOUR %in% c(-500, -100))

ggplot() +
  # Blocks
  geom_sf(data=blocks %>% 
            filter(!(block_type == "Offshore") & block_id < 900), 
          mapping=aes(alpha=mpa_yn), lwd=0.2) +
  # Contours
  geom_sf(data = contours_500, lwd = 0.2) +
  # MPAS
  geom_sf(data = mpas,
          mapping=aes(fill = type), alpha = 0.5) +
  # Region lines
  geom_hline(mapping=aes(yintercept=region_lats)) +
  # Land
  geom_sf(data=foreign, fill="grey80", color="white", lwd=0.3) +
  geom_sf(data=usa, fill="grey80", color="white", lwd=0.3) +
  # Legend
  scale_fill_discrete(name="MPA type") +
  scale_alpha_manual(name="Block type", values=c(1, 0.2)) +
  # Crop
  coord_sf(xlim = c(-122.75, -121.4), ylim = c(36, 37.2)) +
  # Theme
  theme_bw() + my_theme 
```

