---
title: "Testing with Rockfish"
author: "Cori Lopazanski"
date: "2023-08-01"
output: 
  html_document: 
    fig_width: 6
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r}
# Clear workspace
rm(list = ls())

# Packages
library(tidyverse)
library(janitor)
library(sf)
library(rfishbase)

# Directories
basedir <- "/Users/lopazanski/Library/CloudStorage/GoogleDrive-lopazanski@ucsb.edu/Shared drives/NCEAS MPA network assessment/MPA Network Assessment: Working Group Shared Folder/data/sync-data"
gisdir <- file.path(basedir, "gis_data/processed")
datadir <- "analyses/2performance_fisheries/analyses/blocks"
fishdir <- "/Users/lopazanski/Documents/github/nceas/CDFW-fishing-data"

```

# Read Data

```{r}
mpas <- readRDS(file.path(gisdir, "CA_MPA_polygons.Rds"))
blocks <- readRDS(file.path(datadir, "blocks_by_mlpa_region_w_mpa_stats.Rds"))
landings_raw <- readRDS(file.path(fishdir, "CDFW_2000_2020_landings_receipts.Rds"))
#sp_key <- readRDS(file.path(fishdir, "CDFW_species_key.Rds"))# This seems incomplete
sp_key <- read_csv(file.path(fishdir, "CDFW_species_key.csv"))

```

# Clean Data

## Create rockfish species list

```{r}
gf_rock <- sp_key %>% filter(grepl("rockfish", comm_name))
```

## Clean up landings data

```{r}
landings <- landings_raw %>% 
  # remove invalid blocks
  filter(!(block_type %in% c("Invalid"))) %>% 
  # filter only rockfish species
  filter(species_id %in% gf_rock$spp_code_num)
```

## Create simple blocks dataframe

```{r}
blocks_simple <- blocks %>% 
  sf::st_drop_geometry() %>% 
  select(block_id, mlpa_region, mpa_yn)
```

# Build Data

We define CPUE as in Richerson et al. 2020 (<https://www.sciencedirect.com/science/article/pii/S016578362030045X>). CPUE is weight per [species or species group] per trip, where a trip is the total landings of the [species or species group] landed by a single vessel on the same date in the same port.

Calculate aggregate catch for each individual vessel (i) in each block (j) using a specific gear (k) on a particular day (t):

```{r}
data <- landings %>% 
  # Vessel - Block - Gear - Day
  group_by(vessel_id, block_id, gear_type, gear_id, gear, year, month, date) %>% 
  # Calculate total landings for each
  summarize(total_lbs = sum(landings_lb, na.rm = T)) %>% 
  ungroup() 

test_multiple_blocks <- data %>% 
  group_by(vessel_id, gear_id, gear, year, month, date) %>% 
  summarize(total_lbs = sum(total_lbs, na.rm = T),
            n_blocks = n())

length(test_multiple_blocks$n_blocks[test_multiple_blocks$n_blocks>1]) # 1743 obs > 1 block
1743/132700

test_multiple_gears <- data %>% 
  group_by(vessel_id, year, month, date) %>% 
  summarize(n_gears = length(unique(gear_id)))

length(unique(data$gear_id)) # 48 different gear configurations

length(test_multiple_gears$n_gears[test_multiple_gears$n_gears>1]) #1436 gears > 1 on single date
1436/131253

# Examine most common gear types
gear_types <- data %>% 
  group_by(gear_type, gear_id, gear) %>% 
  summarize(receipts = n(), 
            landings = sum(total_lbs, na.rm = T),
            vessels = length(unique(vessel_id)))

# Plot common gears by year
gear_landings <- data %>% 
  group_by(gear_type, year) %>% 
  summarize(landings = sum(total_lbs, na.rm = T)) 

ggplot(data = gear_landings) + 
  geom_line(aes(x = year, y = landings, color = fct_reorder(gear_type, desc(landings)))) +
  labs(color = NULL, 
       x = "Year", y = "Total rockfish catch (lbs) per major gear type") +
  theme_minimal()

```

Create dummy variables for whether or not there is an MPA in the region on a particular day (t):

```{r}
data <- data %>% 
  # Join the MLPA region and MPA status
  left_join(., blocks_simple) %>% 
  # Create dummy for whether or not there is an MPA in the region for each block
  mutate(mpa_in_region = case_when(date < '2007-09-01' & mlpa_region == "Central Coast" ~ 0,
                                 date < '2010-05-01' & mlpa_region == "North Central Coast" ~ 0,
                                 date < '2012-12-01' & mlpa_region == "North Coast" ~ 0,
                                 date < '2012-01-01' & mlpa_region == "South Coast" ~ 0,
                                 mlpa_region == "San Francicso Bay" ~ 0,
                                 TRUE ~ 1))
```

Create dummy variable for whether or not there is an MPA in the block on a particular day (t):

```{r}
data <- data %>% 
  # Create dummy for whether or not there is an MPA (implemented) in the block
  mutate(mpa_in_block = case_when(mpa_yn == "MPA" & mpa_in_region == 1 ~ 1,
                                  TRUE ~ 0))
```

## Calculate Standardized CPUE

Scale the CPUE to center around the mean and convert to unit variance based on the pre-implementation moments.

Since the network was implemented regionally, which pre-implementation mean do we want to use? One that is consistent before any MPA was implemented? Region-specific?

```{r}
# Start with calculating pre-implementation mean based on zero MLPA MPAs implemented
# e.g. prior to first region implemented (before 09-01-2007; Central)
# which is technically after Channel Islands MPAs and some other MPAs...
data_pre <- data  %>% 
  filter(date < '2007-09-01')

# Calculate mean and standard deviation
data_pre.mean <- mean(data_pre$total_lbs)
data_pre.sd <- sd(data_pre$total_lbs)

# Create new cpue_std = (cpue - mean before impl.)/sd before impl.
data$cpue_std <- (data$total_lbs - data_pre.mean)/data_pre.sd

```

## Difference Pre/Post

```{r}
# Calculate the total standardized CPUE for each block pre-implementation
plot_pre <- data %>% 
  filter(date < '2007-09-01') %>% 
  group_by(block_id) %>% 
  summarize(pre_cpue = sum(cpue_std))

plot_post <- data %>% 
  filter(date > '2007-09-01') %>% 
  group_by(block_id) %>% 
  summarize(post_cpue = sum(cpue_std))

plot <- full_join(plot_pre, plot_post) 

plot$diff <- plot$post_cpue-plot$pre_cpue
```

## Version with Inshore Only

```{r}
# Inshore blocks only
inshore <- blocks %>% 
  filter(block_type != "Offshore")

# Subset data to only the inshore blocks
data_inshore <- data %>% 
  filter(block_id %in% inshore$block_id)


data_inshore_pre <- data_inshore %>% 
  filter(date < '2007-09-01')
  
data_inshore_pre.mean <- mean(data_inshore_pre$total_lbs)
data_inshore_pre.sd <- sd(data_inshore_pre$total_lbs)
data_inshore$cpue_std <- (data_inshore$total_lbs - data_inshore_pre.mean)/data_inshore_pre.sd
```

## Difference Pre/Post

```{r}
inshore_pre <-data_inshore %>% 
  filter(date < '2007-09-01') %>% 
  group_by(block_id) %>% 
  summarize(pre_cpue = sum(cpue_std))

inshore_post <- data_inshore %>% 
  filter(date > '2007-09-01') %>% 
  group_by(block_id) %>% 
  summarize(post_cpue = sum(cpue_std))

plot_inshore <- full_join(inshore_pre, inshore_post) 

plot_inshore$diff <- plot_inshore$post_cpue-plot_inshore$pre_cpue

```

# Plots

## Theme Setup

```{r}
# CA/OR, Alder Creek, Pigeon Point, Point Conception, CA/MEX 
region_lats <- c(42, 39.0, 37.18, 34.5, 32.5) %>% rev()

# Get land
usa <- rnaturalearth::ne_states(country="United States of America", returnclass = "sf")
foreign <- rnaturalearth::ne_countries(country=c("Canada", "Mexico"), returnclass = "sf")

# Theme
my_theme <-  theme(axis.text=element_text(size=7),
                   axis.text.y = element_text(angle = 90, hjust = 0.5),
                   axis.title=element_blank(),
                   legend.text=element_text(size=8),
                   legend.title=element_text(size=9),
                   # Gridlines
                   panel.grid.major = element_blank(), 
                   panel.grid.minor = element_blank(),
                   panel.background = element_blank(), 
                   axis.line = element_line(colour = "black"),
                   # Legend
                   legend.key.size = unit(0.4, "cm"),
                   legend.key = element_rect(fill=alpha('blue', 0)),
                   legend.background = element_rect(fill=alpha('blue', 0)),
                   legend.position = c(0.7,0.8))
```

## All Blocks

```{r}
ggplot() +
  # Blocks
  geom_sf(data = blocks, lwd = 0.2) +
  # Landings 
  geom_sf(data = plot %>% 
            left_join(., blocks), aes(fill = diff, geometry = geometry)) +
  # MPAs
  geom_sf(data = mpas, color = "black", lwd = 0.2) +
  # Region lines
  geom_hline(mapping=aes(yintercept=region_lats)) +
  # Land
  geom_sf(data=foreign, fill="grey80", color="white", lwd=0.3) +
  geom_sf(data=usa, fill="grey80", color="white", lwd=0.3) +
  # Crop
  coord_sf(xlim = c(-126, -117), ylim = c(32.5, 42)) +
  # Color
  scale_fill_gradient2(midpoint = 0) +
  theme_bw() + my_theme + 
  labs(title = "Rockfish (Sebastes spp.)",
       fill = "Difference in std. CPUE")

ggsave("rockfish_all_diff.png", width = 4, height = 6, units = c("in"), dpi = 150)
```

## Inshore Blocks

```{r}
ggplot() +
  # Blocks
  geom_sf(data = blocks, lwd = 0.2) +
  # Landings 
  geom_sf(data = plot_inshore %>% 
            left_join(., blocks), aes(fill = diff, geometry = geometry)) +
   # MPAs
  geom_sf(data = mpas, color = "black", lwd = 0.2) +
  # Region lines
  geom_hline(mapping=aes(yintercept=region_lats)) +
  # Land
  geom_sf(data=foreign, fill="grey80", color="white", lwd=0.3) +
  geom_sf(data=usa, fill="grey80", color="white", lwd=0.3) +
  # Crop
  coord_sf(xlim = c(-126, -117), ylim = c(32.5, 42)) +
  # Color
  scale_fill_gradient2(midpoint = 0) +
  theme_bw() + my_theme + 
  labs(title = "Rockfish (Sebastes spp.)",
       fill = "Standardized difference post-MPA")

```
